---
title: "KODAQS Academy | Course D2 Survey"
subtitle: "*Assignment Task Group USA*"

date: today
date-format: "DD.MM.YYYY"
date-modified	: last-modified
author:
  - name: Josefina Sommer
  - name: Elnari Potgieter
  - name: David Simon
  - name: Ana Cristancho
  - name: Lukas Rädle
  - name: Erik Paessler
format:
  html:
    echo: true
    warning: false
    message: false 
    code-overflow: scroll
    code-copy: true
    code-fold: false
    toc: true
    toc-location: left
    theme: cosmo
    html-math-method: katex
---



## Introduction (ALL)

This is an R-Quarto template which can be used for your final assignment task of KODAQS Academy, Course D2 Survey. [How to create citable article](https://quarto.org/docs/authoring/create-citeable-articles.html)<br>
It contains prepared sections for each kind of data quality indicators. Choose an appropriate—for data quality related research questions—instrument from the construct ***Trust*** or ***Attitude***.
For each data quality indicator, you have to select indices that are part of the dataset and/or can be calculated, and specify whether and how you will flag/filter cases. 
Our final assignment task contains parts that should be worked on in a collaborative way (ALL) and parts that should worked on by individual team members (YOUR NAME).
Each teammate should be responsible for at least one indicator section.



::: {.callout-caution appearance="default"}

## Pay Attention Please!
For our last WT prepare a short ppt-presentation (5-10 min max.) where you outline
your analysis strategy (i.e. identified DQIs, and flagging/filtering rules) and show first findings/results. The preparation for and presention/participation on the weekly touchpoint is mandatory and weighs one third of your individual assignment grade.
**WEEKLY TOUCHPOINT (mandatory) will take place on June 5th, 2025, 11–12 am CET**

:::



## Data Preparation (ALL)

In a first step you have to define your R setup (e.g., install und load relevant packages) and need to read in the dataset. Do your basic cleanup and wrangling here (e.g. filtering for your country group, selection of used variables etc.). 

### Load Packages and Read Data
```{r}
####  ------------- Load Packages ------------- ####
library(here)           # find files
library(tidyverse)      # data management
library(skimr)          # descriptives
library(labelled)       # data labels

####  ------------- Read Data ------------- ####

# There are two versions of the data (!!!):
# 
# data_raw: The raw dataset (N = 167,101) before any cleaning and transformations.
#
# data: The analysis-ready dataset (N = 69,534), including post-stratification weights:https://osf.io/5c3qd/wiki?wiki=fqmdn

# data_raw
data_raw <- read.csv(here("00_data/ds_full.csv"), sep = ";") %>%
  filter(COUNTRY_CODE == "USA")  %>%
  mutate(across(where(is.numeric), ~ na_if(., 99))) %>%
  mutate(across(where(is.character), ~ na_if(., "99")))

# raw daat needs to be processed (a liitle bit)
# Source: https://osf.io/5c3qd/files/xz8bw

# (1.)  Exclude incomplete records (all respondents who did not reach the last page)
data_raw_tmp <- data_raw %>%
  filter(!is.na(DEM_RESIDENCE)) 

# % removed
per_rem1 <- ((nrow(data_raw) - nrow(data_raw_tmp)) / nrow(data_raw)) * 100

data_raw <- data_raw_tmp
rm(data_raw_tmp)

# (2.) Identify and exclude duplicate completes

# Initialize ID data (stored in separate csv file)
ds_id <- read.csv(here("00_data/ds_id.csv"), sep = ";") 

data_raw_tmp <- left_join(data_raw, ds_id, by = "ID_QUALTRICS") %>%
  group_by(ID_POLLCMPNY) %>%
  filter(is.na(ID_POLLCMPNY) | row_number(Recorded) == 1) %>%
  ungroup()

# Confirm that there are no duplicate completes (empty tibble)
data_raw_tmp %>% 
  filter(!is.na(ID_POLLCMPNY)) %>%
  group_by(ID_POLLCMPNY) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)

# % removed
per_rem2 <- ((nrow(data_raw) - nrow(data_raw_tmp)) / nrow(data_raw)) * 100

data_raw <- data_raw_tmp
rm(ds_id)
rm(data_raw_tmp)

# (3.) Exclude outlier values for age
data_raw <- data_raw %>%
  mutate(
    DEM_AGE = case_when(
      DEM_AGE < 18 | DEM_AGE > 100 ~ NA_real_,
      TRUE ~ DEM_AGE
    )
  )

# % outliers
per_out <- mean(!(data_raw$DEM_AGE %in% 18:100) & !is.na(data_raw$DEM_AGE)) * 100

# data
data <- read.csv(here("00_data/ds_final.csv"), sep = ";") %>%
  filter(COUNTRY_CODE == "USA")
```

The following packages are used in this script.
```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

Two datasets are used:

- `data_raw`: The raw dataset (N = 167,101) before any cleaning or transformations.
- `data`: The analysis-ready dataset (N = 69,534), including post-stratification.

Not all calculations can be performed using the pre-processed data, as some information is already missing. Therefore, the raw data were processed as follows:

1. Respondents who did not reach the last page were excluded: `r round(per_rem1, 2)`%
2. Duplicate respondents were excluded: `r round(per_rem2, 2)`%
3. Age values outside the range of 18 to 100 were set to NA: `r round(per_out, 2)`%
4. Gender was temporarily recoded into a binary variable (1 = male), with self-described responses excluded from this coding.
5. tbc

*Note to the group: I think most calculations should be done with the file 'data_raw'.*

### Compute Indices

*Science-related populist attitudes* were measuered with 8 items on 4 dimensions on a 5-point scale each (1 = strongly disagree to 5 = strongly agree; Mede, Schäfer, & Füchslin, 2021), aggregated with the Goertz approach (Wuttke et al., 2020)

```{r}
####  ------------ Compute Indices ------------ ####

# SCIPOP_goertz: Science-related populist attitudes

# Compute overall and dimension scores

data_raw <- data_raw %>%
  mutate(
    SCIPOP_PPL_m = rowMeans(across(c(SCIPOP_common, SCIPOP_good)), na.rm = TRUE),
    SCIPOP_ELI_m = rowMeans(across(c(SCIPOP_advantage, SCIPOP_cahoots)), na.rm = TRUE),
    SCIPOP_DEC_m = rowMeans(across(c(SCIPOP_influence, SCIPOP_involved)), na.rm = TRUE),
    SCIPOP_TRU_m = rowMeans(across(c(SCIPOP_lifeexp, SCIPOP_rely)), na.rm = TRUE),
    SCIPOP_goertz = pmin(SCIPOP_PPL_m, SCIPOP_ELI_m, SCIPOP_DEC_m, SCIPOP_TRU_m, na.rm = TRUE)
  )

data_raw <- labelled::set_variable_labels(
  data_raw,
  SCIPOP_goertz = "Science-related populist attitudes: overall (Goertz score)",
  SCIPOP_PPL_m = "Science-related populist attitudes: conceptions of the ordinary people (mean)",
  SCIPOP_ELI_m = "Science-related populist attitudes: conceptions of the academic elite (mean)",
  SCIPOP_DEC_m = "Science-related populist attitudes: demands for decision-making sovereignty (mean)",
  SCIPOP_TRU_m = "Science-related populist attitudes: demands for truth-speaking sovereignty (mean)"
)

# Gender (1 = male)
data_raw <- data_raw  %>%
  mutate(DEM_GENDER = ifelse(DEM_GENDER == "Male", 1, 0)) %>%
  mutate(DEM_GENDER = as.factor(DEM_GENDER))
```

## Descriptive Analysis (ALL)

Provide basic descriptive statistics for your (sub)sample and instrument(s).

```{r}
data_des <- data_raw  %>%
  select(DEM_AGE, DEM_GENDER, # demographics
         SCIPOP_common, SCIPOP_good, SCIPOP_advantage, SCIPOP_cahoots, SCIPOP_influence, SCIPOP_involved, SCIPOP_lifeexp, SCIPOP_rely,# items
         SCIPOP_PPL_m, SCIPOP_ELI_m, SCIPOP_DEC_m, SCIPOP_TRU_m, SCIPOP_goertz) %>% # scales
  skim()

data_des
```


## Attention Checks (David)

The first kind of data quality indicators are check items. What check items are provided in the dataset and did participants pass the check?

```{r}
attcheck1.fails <- function(attcheck1){
  n_valid <- sum(!is.na(attcheck1), na.rm = T)
  n_fails <- sum(attcheck1 != "213", na.rm = T)
  return((1 - ((n_valid - n_fails) / n_valid)) * 100)
}


attcheck2.fails <- function(attcheck2){
  n_valid <- sum(!is.na(attcheck2), na.rm = T)
  n_fails <- sum(attcheck2 %in% 2:5, na.rm = T) + sum(attcheck2 == -99, na.rm = T)
  return((1 - ((n_valid - n_fails) / n_valid)) * 100)
}
```

The first attention of writing “213” into a text box was failed by: `r round(attcheck1.fails(data_raw$ATTCHECK_NUMBER), 2)`% of participants.

`r round(attcheck2.fails(data_raw$ATTCHECK_RES), 2)`%of participants who reached the second attention check of selecting “strongly disagree” for an extra item in a scale of science-related populist attitudes did not pass it.

## Para Data (YOUR NAME)

The second kind of data quality indicators are para data. What kind of para data can you find in the data? Show with an example how this type of variable can be used for DQA! 

```{r}



```



## Missing Data (YOUR NAME)

The third kind of data quality indicators are missing data. How much data is missing, which data missing pattern occurs and are there cases with many missing data? (in respect to your choosen instrument(s) / variable(s)).

```{r}



```



## Response Distributions (YOUR NAME)

The fourth kind of data quality indicators are response distributions. Which indicators do you deem relevant and which criterion do you use for flagging?

```{r}



```



## Response Styles (YOUR NAME)

The last kind of data quality indicators are response styles. Which style(s) do you want to focus on and how do you flag concerning cases?

```{r}



```

## All together (ALL)

Now that you have different flags you may want to combine them. Which combinations do you deem suitable? And how to (partly) filter the identified cases?

```{r}



```

## Discuss (ALL)

Discuss central findings shortly! Make some statements on your final decision in respect to include/exclude rules (cutoffs/thresholds/filter combination), data loss, consequences for planned analyses, and any notable what you like to adress.


## References (ALL)

Place citable refs (used packages and literature) here.
[How to add appendices](https://quarto.org/docs/authoring/appendices.html)


::: {.callout-important appearance="default"}

## Pay Attention Please!
For final submission upload your prepared Quarto document and an exported html-file 
into Moodle, week 8, section "Final Assignment Course D2 Survey LU8.1".
**FINAL DEADLINE will be June 20th, 2025, sharp!**

:::

<span style="color:#1E8CC8;">**Good luck and see you next Thursday (June 5th) on our last weekly touchpoint!**</span>👋



