---
title: "KODAQS Academy | Course D2 Survey"
subtitle: "*Assignment Task Group USA*"

date: today
date-format: "DD.MM.YYYY"
date-modified	: last-modified
author:
  - name: Josefina Sommer
  - name: Elnari Potgieter
  - name: David Simon
  - name: Ana Cristancho
  - name: Erik Paessler
format:
  html:
    echo: true
    warning: false
    message: false 
    code-overflow: scroll
    code-copy: true
    code-fold: false
    toc: true
    toc-location: left
    theme: cosmo
    html-math-method: katex
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
bibliography: references.bib
---


## Introduction (ALL)

This assignment explores data quality aspects of United States of America (USA) data set of the Trust in Science and Science-related Populism (TISP) project. The TISP project consists of a consortium of 241 researchers from 179 institutions focused on global public opinion regarding science. Data was collected from 68 countries from November 2022 to August 2023, with the survey gathering the opinions of 71,922 respondents in total (TISP, n.d).

The project measures various constructs and characteristics. We focus on the questionnaire items related to science-related populism for this assignment.

*Science-related populism* has been conceptualized as a set of ideas suggesting a fundamental conflict between an allegedly virtuous people and an allegedly immoral academic elite over who should be in charge of science-related decision-making and over what is deemed true knowledge [@mede2020].

Conceptually, science-related populism consists of four major components: 

+ the ordinary people, 
+ the academic elite, 
+ science-related decision-making sovereignty, and 
+ truth-speaking sovereignty.

This is the  final assignment task of KODAQS Academy, Course D2 Survey of "Team USA".

As our interest here is in the data quality aspects of the dataset, we consider the findings of various data quality tests - including attention checks, para data, missing data, response distribution and response styles – to discern the overall quality of the USA dataset and to highlight any respondents that should be flagged due to their response behaviours. 

Some parts of the assignment were intended to be completed collaboratively and others individually, each teammate was responsible for at least one section of the indicator. Overall, we worked together constructively and collaboratively on this script. Consequently, the listed names do not fully capture the collaborative nature of the work, as all team members contributed across multiple sections.

## Data Preparation (ALL)

In a first step you have to define your R setup (e.g., install und load relevant packages) and need to read in the dataset. Do your basic cleanup and wrangling here (e.g. filtering for your country group, selection of used variables etc.). 

### Load Packages and Read Data
```{r}
####  ------------- Load Packages ------------- ####
library(here)           
library(tidyverse)     
library(skimr)          
library(labelled)       
library(psych)
library(stringr)
library(resquin)
library(gt)
library(ggcorrplot)
library(visdat)
library(tibble)
library(naniar)

####  ------------- Read Data ------------- ####

# data_raw
data_raw <- read.csv(here("00_data/ds_full.csv"), sep = ";") %>%
  filter(COUNTRY_CODE == "USA")  %>%
  mutate(across(where(is.numeric), ~ na_if(., 99))) %>%
  mutate(across(where(is.character), ~ na_if(., "99")))

# raw data needs to be processed
# Source: https://osf.io/5c3qd/files/xz8bw

# (1.) Identify and exclude duplicate completes

# Initialize ID data (stored in separate csv file)
ds_id <- read.csv(here("00_data/ds_id.csv"), sep = ";") 

data_raw_tmp <- left_join(data_raw, ds_id, by = "ID_QUALTRICS") %>%
  group_by(ID_POLLCMPNY) %>%
  filter(is.na(ID_POLLCMPNY) | row_number(Recorded) == 1) %>%
  ungroup()

# Confirm that there are no duplicate completes (empty tibble)
data_raw_tmp %>% 
  filter(!is.na(ID_POLLCMPNY)) %>%
  group_by(ID_POLLCMPNY) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)

# % removed
per_rem1 <- ((nrow(data_raw) - nrow(data_raw_tmp)) / nrow(data_raw)) * 100

data_raw <- data_raw_tmp
rm(ds_id)
rm(data_raw_tmp)

# (2.) Exclude outlier values for age
# % outliers
per_out <- mean(data_raw$DEM_AGE < 18 | data_raw$DEM_AGE > 100, na.rm = TRUE) * 100

# Set to NA
data_raw <- data_raw %>%
  mutate(
    DEM_AGE = if_else(DEM_AGE < 18 | DEM_AGE > 100, 
                      as.numeric(NA),
                      DEM_AGE)         
  )

# (3.) Gender (1 = male)
data_raw <- data_raw  %>%
  mutate(DEM_GENDER = ifelse(DEM_GENDER == 1, 1, 0)) %>%
  mutate(DEM_GENDER = as.factor(DEM_GENDER))

# (4.) Exclude respondents who did not reach the first question = DEM_GENDER
data_raw_tmp <- data_raw %>%
  filter(!is.na(DEM_GENDER))

# % removed
per_rem2 <- ((nrow(data_raw) - nrow(data_raw_tmp)) / nrow(data_raw)) * 100

data_raw <- data_raw_tmp
rm(data_raw_tmp)
```

The following packages are used in this script:

```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

One dataset is read:

- `ds_full.csv`: The raw dataset before any cleaning or transformations.

The raw data was processed as follows:

1. The data was filtered using `COUNTRY_CODE == "USA"`
2. Duplicate respondents were excluded: `r round(per_rem1, 2)`%
3. Age values outside the range of 18 to 100 were set to NA: `r round(per_out, 2)`%
4. Gender was recoded into a binary variable (1 = male), with self-described responses excluded from this coding.
5. Only participants were included who reached the beginning of the actual questionnaire (= question 1: DEM_GENDER): `r round(per_rem2, 2)`%

We performed all calculations on this dataset `data_raw`.

### Compute Indices

*Science-related populist attitudes* were measuered with 8 items on 4 dimensions on a 5-point scale each (1 = strongly disagree to 5 = strongly agree) [@mede2020], aggregated with the Goertz approach [@wuttke2020]. The script was privided by Mede et al. [-@mede2025].

```{r}
# SCIPOP_goertz: Science-related populist attitudes

# Compute overall and dimension scores

data_raw <- data_raw %>%
  mutate(
    SCIPOP_PPL_m = rowMeans(across(c(SCIPOP_common, SCIPOP_good)), na.rm = TRUE),
    SCIPOP_ELI_m = rowMeans(across(c(SCIPOP_advantage, SCIPOP_cahoots)), na.rm = TRUE),
    SCIPOP_DEC_m = rowMeans(across(c(SCIPOP_influence, SCIPOP_involved)), na.rm = TRUE),
    SCIPOP_TRU_m = rowMeans(across(c(SCIPOP_lifeexp, SCIPOP_rely)), na.rm = TRUE),
    SCIPOP_goertz = pmin(SCIPOP_PPL_m, SCIPOP_ELI_m, SCIPOP_DEC_m, SCIPOP_TRU_m, na.rm = TRUE)
  )

data_raw <- labelled::set_variable_labels(
  data_raw,
  SCIPOP_goertz = "Science-related populist attitudes: overall (Goertz score)",
  SCIPOP_PPL_m = "Science-related populist attitudes: conceptions of the ordinary people (mean)",
  SCIPOP_ELI_m = "Science-related populist attitudes: conceptions of the academic elite (mean)",
  SCIPOP_DEC_m = "Science-related populist attitudes: demands for decision-making sovereignty (mean)",
  SCIPOP_TRU_m = "Science-related populist attitudes: demands for truth-speaking sovereignty (mean)"
)
```


## Descriptive Analysis (ALL)

Provide basic descriptive statistics for your (sub)sample and instrument(s).

```{r}
data_des <- data_raw  %>%
  select(DEM_AGE, DEM_GENDER, # demographics
         SCIPOP_common, SCIPOP_good, SCIPOP_advantage, SCIPOP_cahoots, SCIPOP_influence, SCIPOP_involved, SCIPOP_lifeexp, SCIPOP_rely,# items
         SCIPOP_PPL_m, SCIPOP_ELI_m, SCIPOP_DEC_m, SCIPOP_TRU_m, SCIPOP_goertz) %>% # scales
  skim()

data_des
```


## Attention Checks (David Simon)

The first kind of data quality indicators are check items. What check items are provided in the dataset and did participants pass the check?

```{r}
attcheck1.fails <- function(attcheck1){
  n_valid <- sum(!is.na(attcheck1), na.rm = T)
  n_fails <- sum(attcheck1 != "213", na.rm = T)
  return((1 - ((n_valid - n_fails) / n_valid)) * 100)
}


attcheck2.fails <- function(attcheck2){
  n_valid <- sum(!is.na(attcheck2), na.rm = T)
  n_fails <- sum(attcheck2 %in% 2:5, na.rm = T) + sum(attcheck2 == -99, na.rm = T)
  return((1 - ((n_valid - n_fails) / n_valid)) * 100)
}
```

Using the provided code by Mede et al. [-@mede2025], the first attention of writing “213” into a text box was failed by `r round(attcheck1.fails(data_raw$ATTCHECK_NUMBER), 0)`% of participants and

`r round(attcheck2.fails(data_raw$ATTCHECK_RES), 0)`% of participants who reached the second attention check of selecting “strongly disagree” for an extra item in a scale of science-related populist attitudes did not pass it.

So, the data set is partially filtered by design: Participants who failed the first attention check were redirected to the end of the survey and did not participate further. Consequently, they are not included among those who answered the scale of science-related populist attitudes.

The second attention check was part of the scale of science-related populist attitudes and is therefore particularly interesting for this assignment. Participants who failed the second attention check were also redirected to the end of the survey. The attention check was the last item on the scale. 

Several participants answered the questions incorrectly and were flagged as having failed by the authors. One could argue about whether these were truly careless respondents. This is particularly questionable for those who wrote the correct number in words instead of as a number, e.g., "two hundred thirteen.", one person used a lowercase L instead of 1. Others are more likely to be careless, e.g. "Very nice to have a good time at your new office and I love love you bye bye good mama bye mama good night too love love you". Others stay unclear or "missing", e.g. "i don't know". The following table provides an overview about the frequency: 

```{r}
attentioncheck_clean <- data_raw$ATTCHECK_NUMBER %>%
  str_to_lower() %>%                       # all lowercase
  str_trim() %>%                           # remove spaces
  str_replace_all("’", "'") %>%            # remove apostrophes
  str_replace_all("[[:punct:]]", " ") %>%  # remove punctuation characters
  str_squish()                             # remove extra spaces

attentioncheck_table <- case_when(
  
    # missing
  is.na(attentioncheck_clean) | attentioncheck_clean %in% c("", "none", "i don t know") | attentioncheck_clean == "99"
  ~ "Missing",
  
  # numeric variants
  str_detect(attentioncheck_clean, "\\b213\\b|\\b2l3\\b") 
  ~ "Correct",
  
  # numeric split with extra spaces
  str_detect(attentioncheck_clean, "^2\\s*1\\s*3$") 
  ~ "Correct",
  
  # word variants
  str_detect(attentioncheck_clean,
             "two.*hund.*thirt|two.*thirt|two.*one.*three|twoh.*thirt|twothirt|twoonethree|doscientos trece") 
  ~ "Correct: written in words",
  
  # incorrect
  TRUE ~ "Incorrect"
)

# add flag to dataset
data_raw <- data_raw %>%
  mutate(
    attentioncheck_flag = case_when(
      attentioncheck_table == "Incorrect" ~ TRUE,
      attentioncheck_table == "Missing" ~ NA,
      TRUE ~ FALSE
    )
  )

table <- table(attentioncheck_table)
prop <- prop.table(table(attentioncheck_table))
table
```
The number of failures on the second attention check decreases from `r round(attcheck2.fails(data_raw$ATTCHECK_RES), 0)`% to `r round(prop.table(prop)["Incorrect"] * 100, 0)`%.

These are candidates of careless respondents and therefore "flagged" in this section.

## Para Data (Elnari Potgieter)

The second kind of data quality indicators are para data. What kind of para data can you find in the data? Show with an example how this type of variable can be used for DQA!

1. Total Response Time

Total response time refers to the time it takes a respondent to complete the survey. The data set contains a "Duration" variable, analysed here, capturing response time in seconds. 

```{r}
# Key Descriptives of Duration variable (in seconds)
describe (data_raw$Duration, na.rm = TRUE)
```
Mean (1090.11) larger than median (620), showing right / positive skew in data. Minimum at 10 seconds, and maximum at 213592 seconds. Further analysis to look out for very high value outliers.

Further consideration to percentiles / quartiles.

```{r}
# Consider quartiles
skim(data_raw$Duration)
# Shows missing value, so proceed and exclude missing values
# Exclude NA values and skim the Duration variable
skim(data_raw$Duration[!is.na(data_raw$Duration)])
```
Maximum/p100 (213592) much higher than mean/p50 (620) and p75 (1101). The extreme high outliers may be due to respondents opening the survey and only answering the survey later. This does not necessarily mean they were not paying attention when they were answering, so I don't think their answers are necessarily invalid. However, on the lower end, 10 seconds for a survey with 111 items (Mede et al., 2025) - some of which open ended - sounds very short. So too 283 for p25. The extreme outliers also impact further calculations.

Box plots (in seconds) for further investigation.

```{r}
# Boxplot excluding NA values
boxplot(data_raw$Duration[!is.na(data_raw$Duration)],
        main = "Boxplot of Durations",
        ylab = "Duration (seconds)",
        col = "lightblue",
        border = "darkblue")
```
Boxplot reflects very high value outliers, which is less of concern in terms of response quality than the very low value durations as the former may still have paid attention, but they do impact the mean and median, which impacts further analysis. I considered running further analysis without the 1st and99th percentile respondents, but after consulting the course example and Appendix B of [-@leiner2019], it appears that extreme outliers are not necessarily removed at this point. 

2. Average Time per Item

The average time per item indicator captures the average (mean) time spent on survey questions.

```{r}
# Number of items
num_items <- 111

# Calculate the average time per item, excluding NA values in Duration
average_time_per_item <- data_raw$Duration/ num_items

# Compute the mean of all average times per item, excluding NA values
mean_average_time_per_item <- mean(average_time_per_item, na.rm = TRUE)

# Output the results
cat("Average time per item for each respondent (in seconds):\n")
cat("\nMean of all average times per item (in seconds):", mean_average_time_per_item, "seconds\n")
```
Using the filtered durations, the average time per item is 9.82 seconds.

I do think it prudent to consider the low value outliers more. I proceed by looking at the number and proportion of respondents with an average time per item below 2 seconds (as per the course example).

```{r}
# Identify respondents who fall below the cutoff of 2 seconds (as per example in course)
cutoff <- 2
below_cutoff <- which(average_time_per_item < cutoff)

# Count of respondents below the cutoff
count_below_cutoff <- length(below_cutoff)

# Proportion of respondents below the cutoff
proportion_below_cutoff <- count_below_cutoff / length(data_raw$Duration)

cat("\nRespondents below the cutoff of", cutoff, "seconds:\n")
cat("\nCount of respondents below the cutoff:", count_below_cutoff, "\n")
cat("Proportion of respondents below the cutoff:", proportion_below_cutoff, "\n")
```
845 respondents (18,37%) had an average time per item below 2 seconds. I am reluctant to exclude respondents only based on this criteria and will proceed with further analysis.

3. Relative Time Completion

Relative time completion compares the speed of completing the questionnaire for a respondent to the median completion time of the sample. 

```{r}
# Median completion time of the sample
median_time <- 620
# As per descriptives

# Calculate the relative completion time for each respondent, excluding NA values
relative_completion_time <- data_raw$Duration / median_time

# Exclude NA values from the relative completion time
relative_completion_time <- relative_completion_time[!is.na(relative_completion_time)]

# Output the relative completion times
cat("Relative completion time for each respondent (excluding NA):\n")

# Summary statistics for relative completion times
cat("\nSummary of relative completion times:\n")
print(summary(relative_completion_time))
```
On the lower end, there are respondents much faster than the median. This aligns with earlier flagged concerns regarding some respondents' fast completion rates. 

As per the course example, we calculate how many respondents are faster than twice the median completion time.

```{r}
# Identify respondents who are faster than two times the median completion time
cutoff <- 2
faster_than_cutoff <- which(relative_completion_time > cutoff)

# Count of respondents faster than two times the median completion time
count_faster <- length(faster_than_cutoff)

# Proportion of respondents faster than two times the median completion time
proportion_faster <- count_faster / length(data_raw$Duration)

# Output the results
cat("\nCount of respondents faster than two times the median:", count_faster, "\n")
cat("Proportion of respondents faster than two times the median:", proportion_faster, "\n")
```
883 (20,24%) respondents where twice as fast as the median of the relative completion time. I am reluctant to exclude this many respondents based on this criteria, so I proceed with also considering the speeding factor as per the course example.

4. Speeding Factor

[@leiner2019] suggests a factor of larger than two as outliers.This is calculated using their response time in relation to the median of the sample. 

```{r}
# Median completion time of the sample
median_time <- 620

# Calculate the speeding factor for each respondent
speeding_factor <- median_time / data_raw$Duration

# Exclude NA values from the speeding factor
speeding_factor <- speeding_factor[!is.na(speeding_factor)]

# Output the speeding factors
cat("Speeding factor for each respondent (excluding NA):\n")

# Summary statistics for speeding factors (excluding NA values)
cat("\nSummary of speeding factors (excluding NA):\n")
print(summary(speeding_factor))
```
From the above summary of speeding factors, the data set does include respondents with a speeding factor above 2.

```{r}
# Identify outliers (speeding factor > 2)
outliers <- which(speeding_factor > 2)

# Count of outliers
outlier_count <- length(outliers)

# Proportion of outliers
outlier_proportion <- outlier_count / length(data_raw$Duration)

# Output the count and proportion of outliers
cat("\nCount of outliers (speeding factor > 2):", outlier_count, "\n")
cat("Proportion of outliers:", outlier_proportion, "\n")
```
The dataset has 1197 (27,44%) - more than a quarter of - respondents with speeding factor above 2.

```{r}
# Add a flag for speeding_factor using case_when
data_raw <- data_raw %>%
  mutate(
    speeding_factor = median(median_time, na.rm = TRUE) / data_raw$Duration,
    paradata_flag = case_when(
      is.na(speeding_factor) ~ NA,
      speeding_factor > 2 ~ TRUE,
      TRUE ~ FALSE
    )
  )

# Set TRUE if speeding_factor > 2
# Set FALSE otherwise
```

5. Summary

With 845 respondents (18,37%) with an average time per item below 2 seconds, 883 (20,24%) respondents twice as fast as the median of the relative completion time and 1197 (27,44%) respondents with speeding factor above 2, it does appear like many respondents were answering the questionnaire quite speedily. 

I ran the same analysis, "trimming" the 1st and 99th percentile to see if the higher value outliers affect the results in a meaningful way, but this was not the case as the mean remains very similar. 

I would be reluctant to exclude respondents based on response times only, as 18-27% of responses would be lost this way. My suggestion would be to align respondents flagged for response times with other tests - for example attention checks, response patterns and response styles.


## Missing Data (Ana)

The third kind of data quality indicators are missing data. How much data is missing, which data missing pattern occurs and are there cases with many missing data? (in respect to your choosen instrument(s) / variable(s)).

```{r}
####  ------------- Load Packages ------------- ####
ds_usa <- read.csv(here("00_data/ds_full.csv"), sep = ";") %>%
  filter(COUNTRY_CODE == "USA")

trust_sci_usa <- ds_usa |>
  dplyr::select(dplyr::starts_with("TRUST_SCI_"))

clim_gov_usa <- ds_usa |>
  dplyr::select(dplyr::starts_with("CLIM_GOV_"))

clim_emo_usa <- ds_usa |>
  dplyr::select(dplyr::starts_with("CLIM_EMO_"))

scipop_usa <- ds_usa |>
  dplyr::select(dplyr::starts_with("SCIPOP_"))

missing_diag_usa <- ds_usa |>
  dplyr::select(
    dplyr::starts_with("TRUST_SCI_"),
    dplyr::starts_with("CLIM_GOV_"),
    dplyr::starts_with("CLIM_EMO_"),
    dplyr::starts_with("SCIPOP_")
  )

```

```{r}
visdat::vis_miss(missing_diag_usa)

```
The missingness map indicates that nonresponse is structured rather than randomly dispersed across items. Missing values cluster in recurring horizontal bands and battery-level blocks, suggesting that missingness is driven primarily by section-level omission, routing, or broader respondent-level nonresponse rather than isolated item skipping.

```{r} 
naniar::gg_miss_var(missing_diag_usa, show_pct = TRUE)

```

The trust-in-scientists battery shows moderate but highly uniform missingness across items, with each item missing in roughly 22% of cases. This suggests that missingness in this battery is not item-specific, but instead reflects a common section-level process affecting all trust items similarly.

```{r}
naniar::gg_miss_upset(missing_diag_usa)
VIM::aggr(missing_diag_usa) |> plot(numbers = TRUE, prop = FALSE)

```

Missingness is concentrated in recurring blocks of variables rather than dispersed randomly across items. This suggests structured battery- or section-level missingness rather than simple random item-level nonresponse. 

#Person-level Missingness

```{r}
n_miss_per_case <- is.na(missing_diag_usa) |>
  as_tibble() |>
  rowSums()

missing_cases_usa <- tibble(
  respondent_id = seq_len(nrow(missing_diag_usa)),
  n_missing = n_miss_per_case
)

summary(missing_cases_usa$n_missing)

```
Respondent-level missingness across the combined batteries was highly uneven. While at least half of respondents had no missing values, the upper quartile was missing 16 or more items, and the maximum was 36 missing items.
```{r}
missing_cases_usa |>
  dplyr::filter(n_missing == ncol(missing_diag_usa)) |>
  print(n = 100)

naniar::gg_miss_case(missing_diag_usa, show_pct = TRUE)

missing_cases_usa |>
  dplyr::filter(n_missing > 0) |>
  dplyr::arrange(dplyr::desc(n_missing)) |>
  print(n = 100)

missing_cases_usa |>
  dplyr::summarise(
    n_with_any_missing = sum(n_missing > 0),
    pct_with_any_missing = mean(n_missing > 0) * 100
  )

missing_cases_usa |>
  dplyr::count(n_missing) |>
  dplyr::arrange(dplyr::desc(n_missing))

ncol(missing_diag_usa)

```


```{r}
battery_missing_summary <- tibble(
  battery = c("TRUST_SCI", "CLIM_GOV", "CLIM_EMO", "SCIPOP"),
  n_items = c(
    ncol(trust_sci_usa),
    ncol(clim_gov_usa),
    ncol(clim_emo_usa),
    ncol(scipop_usa)
  ),
  mean_pct_missing = c(
    mean(is.na(as.matrix(trust_sci_usa))) * 100,
    mean(is.na(as.matrix(clim_gov_usa))) * 100,
    mean(is.na(as.matrix(clim_emo_usa))) * 100,
    mean(is.na(as.matrix(scipop_usa))) * 100
  )
)

battery_missing_summary

```


#Item-Level Missingness

```{r}
item_missing_usa <- missing_diag_usa |>
  dplyr::summarise(
    across(everything(), ~ mean(is.na(.x)) * 100)
  ) |>
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "pct_missing"
  ) |>
  dplyr::arrange(dplyr::desc(pct_missing))

item_missing_usa

ggplot(item_missing_usa, aes(x = reorder(variable, pct_missing), y = pct_missing)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    x = "Variable",
    y = "% missing",
    title = "Item-level missingness in selected USA survey batteries"
  ) +
  theme_minimal()

```

The trust-in-scientists battery shows moderate but very even missingness across items, suggesting that nonresponse reflects a common battery-level process rather than selective skipping of particular trust items.

#Domain-level missingness

```{r}
domain_missing_usa <- ds_usa |>
  dplyr::transmute(
    miss_trust_sci = as.integer(dplyr::if_all(dplyr::starts_with("TRUST_SCI_"), is.na)),
    miss_clim_gov  = as.integer(dplyr::if_all(dplyr::starts_with("CLIM_GOV_"), is.na)),
    miss_clim_emo  = as.integer(dplyr::if_all(dplyr::starts_with("CLIM_EMO_"), is.na)),
    miss_scipop    = as.integer(dplyr::if_all(dplyr::starts_with("SCIPOP_"), is.na))
  )

domain_missing_usa |>
  dplyr::summarise(dplyr::across(everything(), sum))

cor(domain_missing_usa)
round(cor(domain_missing_usa), 3)


```
Battery-level missingness indicators were very strongly correlated, especially between climate-government and climate-emotion items (r = 0.997) and between trust-in-scientists and science-populism items (r = 0.978). This indicates that missingness is driven by structured module-level processes rather than isolated or random nonresponse.

#Complete-case Restriction
```{r}
complete_cases_usa <- missing_diag_usa |>
  tidyr::drop_na()

nrow(missing_diag_usa)
nrow(complete_cases_usa)

tibble(
  total_cases = nrow(missing_diag_usa),
  complete_cases = nrow(complete_cases_usa),
  cases_dropped = nrow(missing_diag_usa) - nrow(complete_cases_usa),
  pct_dropped = (nrow(missing_diag_usa) - nrow(complete_cases_usa)) / nrow(missing_diag_usa) * 100
)

```
Complete-case analysis would exclude 2,067 of 4,662 respondents (44.3%), showing that listwise deletion is highly costly when these batteries are analysed together, , indicating that complete-case analysis would substantially reduce statistical power and is poorly suited to a dataset in which missingness is structured at the battery level.

#Logistic Regression
```{r}

# install.packages("logistf")
library(logistf)


miss_model_trust_sci <- glm(
  miss_trust_sci ~ DEM_AGE + DEM_EDU + Duration + Response_TYPE 
  + DEM_GENDER,
  data = dplyr::bind_cols(ds_usa, domain_missing_usa),
  family = binomial()
)

summary(miss_model_trust_sci)

model_df <- dplyr::bind_cols(ds_usa, domain_missing_usa) |>
  dplyr::filter(
    !is.na(DEM_AGE),
    !is.na(DEM_EDU),
    !is.na(Duration),
    !is.na(Response_TYPE),
    !is.na(DEM_GENDER)
  ) |>
  dplyr::mutate(
    DEM_EDU       = factor(DEM_EDU),
    Response_TYPE = factor(Response_TYPE),
    DEM_GENDER    = factor(DEM_GENDER)
  )

table(model_df$miss_trust_sci, model_df$Response_TYPE)
table(model_df$miss_trust_sci, model_df$DEM_GENDER)
table(model_df$miss_trust_sci, model_df$DEM_EDU)

miss_model_trust_sci <- logistf::logistf(
  miss_trust_sci ~ DEM_AGE + DEM_EDU + Duration + Response_TYPE + DEM_GENDER,
  data = model_df
)
 
summary(miss_model_trust_sci)

tibble::tibble(
  term      = miss_model_trust_sci$terms,
  odds_ratio = exp(coef(miss_model_trust_sci)),
  ci_lower  = exp(miss_model_trust_sci$ci.lower),
  ci_upper  = exp(miss_model_trust_sci$ci.upper),
  p_value   = miss_model_trust_sci$prob
)

```
A penalised logistic regression model was used to account for sparse categories and separation issues. Results indicate that missingness in the trust-in-scientists battery is systematically structured. Older respondents were significantly less likely to exhibit whole-battery missingness, and gender differences were also observed, with one group showing a lower probability of omission. Completion duration was positively associated with missingness, although the effect size was small. In contrast, education was not a robust predictor once modelled categorically. A survey-administration variable (Response Type) exhibited near-perfect separation and was therefore interpreted as a structural artefact rather than a substantive determinant of missingness.

Taken together, these findings suggest that missingness is neither Missing Completely at Random (MCAR) nor purely item-level. Instead, it reflects a combination of respondent characteristics, survey engagement, and design-driven processes operating at the battery level.


## Response Distributions (Erik Paessler)

The fourth kind of data quality indicators are response distributions. Which indicators do you deem relevant and which criterion do you use for flagging?

For response distributions, I focus on **intra-individual (within-person) location and spread indicators** computed with `resquin::resp_distributions()`. Five Likert-scale batteries are examined separately so the statistics remain interpretable within each coherent scale: Trust in Scientists (`TRUST_SCI_`), Climate Government Trust (`CLIM_GOV_`), Climate Emotions (`CLIM_EMO_`), Science Populism (`SCIPOP_`, 8 original items), and Science Information Sources (`SCIINFO_`).

The indicators I deem most relevant are:

- **`ii_sd`** (intra-individual standard deviation): captures within-person response spread. A value of zero flags **straightliners**, i.e. respondents who chose the identical option for every item in a battery, a strong marker of inattentive responding.
- **`mahal`** (Mahalanobis distance): captures how unusual a respondent's response profile is relative to the sample distribution. Respondents in the **bottom 10%** have profiles suspiciously close to the sample centroid (overly uniform, consistent with careless responding), whereas respondents in the **top 10%** show unusually deviant profiles.
- **`ii_mean` / `ii_median`**: secondary indicators used to characterise the nature of flagged profiles (e.g., persistent floor or ceiling responding combined with near-zero SD).

**Flagging criterion**: A case is flagged as a potential data quality concern if it meets at least one of the following for the focal SCIPOP battery: (1) `ii_sd == 0` (straightlining), or (2) Mahalanobis distance in the bottom or top 10% of the sample. All other batteries serve as comparative context.

```{r}
# Extract item batteries from data_raw ----

# Trust in Scientists
trust_sci <- data_raw |> select(starts_with("TRUST_SCI_"))

# Climate Government Trust
clim_gov  <- data_raw |> select(starts_with("CLIM_GOV_"))

# Climate Emotions
clim_emo  <- data_raw |> select(starts_with("CLIM_EMO_"))

# Science Populism: original 8 items only (exclude computed indices added earlier)
scipop <- data_raw |>
  select(SCIPOP_common, SCIPOP_good, SCIPOP_advantage, SCIPOP_cahoots,
         SCIPOP_influence, SCIPOP_involved, SCIPOP_lifeexp, SCIPOP_rely)

# Science Information Sources
sciinfo <- data_raw |> select(starts_with("SCIINFO_"))


# Compute response distribution indicators ----

rd_trust_sci <- resp_distributions(trust_sci)
rd_clim_gov  <- resp_distributions(clim_gov)
rd_clim_emo  <- resp_distributions(clim_emo)
rd_scipop    <- resp_distributions(scipop)
rd_sciinfo   <- resp_distributions(sciinfo)
```

### Descriptive Summaries

```{r}
cat(" Trust in Scientists \n");  summary(rd_trust_sci)
cat(" Climate Gov. Trust \n");   summary(rd_clim_gov)
cat(" Climate Emotions \n");     summary(rd_clim_emo)
cat(" Science Populism \n");     summary(rd_scipop)
cat(" Science Info Sources \n"); summary(rd_sciinfo)
```

The `ii_mean` values differ noticeably across batteries. Trust in Scientists has the highest average mean (3.81 on a 5-point scale), indicating that US respondents generally lean toward trusting scientists. Science Populism and Science Info Sources both sit near 3.44, but on different scales — for SCIPOP (1–5) this is around the midpoint, while SCIINFO appears to use a wider scale. Climate Emotions has the lowest average mean (2.76), suggesting US respondents more often report lower emotional engagement with climate change. Climate Emotions and Climate Government Trust also stand out for missingness: their 75th percentile for `n_na` reaches the maximum possible value, indicating that when respondents have missing items in these batteries, they tend to be missing all items, which indicates structure behind the skip patterns as opposed to random non-response.

### Univariate Distribution Plots

```{r}
plot_univariate <- function(rd, title) {
  rd |>
    select(-contains("na")) |>
    pivot_longer(-id) |>
    ggplot(aes(value)) +
    geom_histogram(bins = 30, fill = "steelblue", color = "white") +
    facet_wrap(vars(name), scales = "free_x") +
    labs(title = title, subtitle = "US respondents", x = NULL, y = "Count") +
    theme_minimal(base_size = 12)
}

plot_univariate(rd_trust_sci, "Response Distributions – Trust in Scientists")
plot_univariate(rd_clim_gov,  "Response Distributions – Climate Gov. Trust")
plot_univariate(rd_clim_emo,  "Response Distributions – Climate Emotions")
plot_univariate(rd_scipop,    "Response Distributions – Science Populism")
plot_univariate(rd_sciinfo,   "Response Distributions – Science Info Sources")
```

The `ii_sd` histograms show a spike at zero across all batteries, indicating straightlining. This spike is most pronounced for **Trust in Scientists** (~650 cases) and **Science Populism** (~625 cases), and notably smaller for Climate Gov. Trust (~350) and Climate Emotions (~300). This is notable because Trust in Scientists, despite expressing the most positive location on average, also contains the most within-person homogeneity cases. The `ii_mean` distributions confirm the substantive differences across batteries: Trust in Scientists is left-skewed (most respondents lean toward 4–5), whereas SCIPOP is roughly symmetric around 3, and Climate Emotions is more evenly spread across the full range with a slight left lean toward lower values. Mahalanobis distances are right-skewed in all batteries, as expected under a chi-squared distribution, with Trust in Scientists showing the most extreme outliers (up to ~10).

### Indicator Boxplots by Battery

```{r}
# Boxplots of each indicator per battery
# (NA-count columns excluded for clarity)

rd_trust_sci |> select(-contains("na")) |> plot(main = "Trust in Scientists – USA")
rd_clim_gov  |> select(-contains("na")) |> plot(main = "Climate Gov. Trust – USA")
rd_clim_emo  |> select(-contains("na")) |> plot(main = "Climate Emotions – USA")
rd_scipop    |> select(-contains("na")) |> plot(main = "Science Populism – USA")
rd_sciinfo   |> select(-contains("na")) |> plot(main = "Science Info Sources – USA")
```

The boxplots complement the histograms by highlighting spread and outlier structure per indicator. Key patterns across batteries:

- **`ii_mean`**: Trust in Scientists has the narrowest IQR positioned high on the scale (IQR ≈ 3.2–4.5), confirming concentrated high-trust responding. Science Info Sources has the widest IQR (≈ 2.1–4.7), reflecting more heterogeneous scale use across its items.
- **`ii_sd`**: Science Info Sources shows the highest median and widest IQR for within-person spread, consistent with its broader scale range. Trust in Scientists and SCIPOP have lower median SD but show many outliers at the upper end, which are respondents who used the full range of the scale.
- **`ii_median`**: SCIPOP has a relatively narrow IQR (≈ 3–4), while Science Info Sources spans nearly the full scale, reflecting its different item content and scale range.
- **`mahal`**: All batteries show heavy right tails with numerous outliers well beyond the whiskers, confirming that a small subset of respondents have somewhat unusual response profiles in each battery. Trust in Scientists produces the most extreme Mahalanobis outliers (up to ~10.2).

### Straightlining: All Batteries

```{r}
# Count respondents with ii_sd == 0 (identical response to every item) in each battery

straight_summary <- tibble(
  Battery    = c("Trust in Scientists", "Climate Gov. Trust", "Climate Emotions",
                 "Science Populism",    "Science Info Sources"),
  n_total    = c(nrow(rd_trust_sci), nrow(rd_clim_gov), nrow(rd_clim_emo),
                 nrow(rd_scipop),    nrow(rd_sciinfo)),
  n_straight = c(
    sum(rd_trust_sci$ii_sd == 0, na.rm = TRUE),
    sum(rd_clim_gov$ii_sd  == 0, na.rm = TRUE),
    sum(rd_clim_emo$ii_sd  == 0, na.rm = TRUE),
    sum(rd_scipop$ii_sd    == 0, na.rm = TRUE),
    sum(rd_sciinfo$ii_sd   == 0, na.rm = TRUE)
  )
) |>
  mutate(pct_straight = round(n_straight / n_total * 100, 1))

print(straight_summary)
```

Straightlining rates are highest for **Trust in Scientists** (15.8%) and **Science Populism** (14.2%), and substantially lower for Climate Gov. Trust (8.0%), Science Info Sources (8.8%), and Climate Emotions (7.1%). The high rate for SCIPOP is of primary concern given its role as the focal construct: roughly 1 in 7 US respondents gave an identical answer to all eight populism items. For Trust in Scientists, the high rate likely reflects the battery's strong positive consensus. Selecting the same high-trust option throughout may reflect overall homogeneity for some respondents rather than inattentiveness, though the two are not distinguishable from response data alone. The lower rates in Climate Emotions may partly reflect that the items tap distinct emotions (e.g., fear, hope, anger), making pure straightlining harder even for inattentive respondents.

### Mahalanobis Distance: All Batteries

```{r}
# Helper function: summarise bottom 10% and top 10% Mahalanobis for a given rd object
mahal_summary <- function(rd, label) {
  p10 <- quantile(rd$mahal, 0.10, na.rm = TRUE)
  p90 <- quantile(rd$mahal, 0.90, na.rm = TRUE)
  cat(sprintf("\n %s \n", label))
  cat(sprintf("  10th pct cutoff: %.3f   |   90th pct cutoff: %.3f\n", p10, p90))
  cat("  Bottom 10% (most uniform profiles):\n")
  print(summary(rd |> filter(mahal <= p10)))
  cat("  Top 10% (most deviant profiles):\n")
  print(summary(rd |> filter(mahal >= p90)))
}

mahal_summary(rd_trust_sci, "Trust in Scientists")
mahal_summary(rd_clim_gov,  "Climate Gov. Trust")
mahal_summary(rd_clim_emo,  "Climate Emotions")
mahal_summary(rd_scipop,    "Science Populism")
mahal_summary(rd_sciinfo,   "Science Info Sources")
```

Across all batteries, the bottom-10% Mahalanobis groups share a consistent signature: near-zero `ii_sd` (often at exactly 0) combined with a modal response location — for Trust in Scientists, the uniform response is near 4–5 (high trust); for SCIPOP, it clusters near 3 (midpoint). This indicates that these are straightlining profiles anchored at the most common category, rather than respondents close to the mean in a multidimensional sense. The SCIPOP bottom 10% is of most interest here: median `ii_sd` = 0 and mean `ii_mean` ≈ 3.25 on a 1–5 scale, a pattern consistent with respondents repeatedly selecting the neutral midpoint without engaging with item differences.

The top-10% groups show higher `ii_sd` and more extreme `ii_mean` values across all batteries. For Trust in Scientists, the top-10% `ii_mean` averages 3.26 with SD of 1.27, which is considerably more variable than the full sample average. These may represent respondents with  nuanced or differentiated views, but given the extreme distances, some may also reflect inattentive responding with random category switches.

### Sensitivity Check: `min_valid_responses` Threshold

```{r}
# CLIM_EMO is used here because it has the highest average number of missing items,
# making it the natural test case for the min_valid_responses threshold.

cat(" CLIM_EMO: default (min_valid = 1.0) \n")
clim_emo |> resp_distributions(min_valid_responses = 1.0) |> summary()

cat(" CLIM_EMO: min_valid = 0.9 \n")
clim_emo |> resp_distributions(min_valid_responses = 0.9) |> summary()

cat(" CLIM_EMO: min_valid = 0.5 \n")
clim_emo |> resp_distributions(min_valid_responses = 0.5) |> summary()
```

The sensitivity analysis shows virtually no change in the summary statistics across all three threshold settings. This is consistent with the pattern already observed in the descriptives: when respondents have missing items in the CLIM_EMO battery, the 75th percentile of `n_na` is at the maximum (9 items), indicating that missings occur in an all-or-nothing fashion. Lowering the threshold therefore does not bring in any partially-completed respondents, because no such partial completions exist in practice. The default strict threshold (1.0) is retained, and `ii_sd` and `mahal` remain reliable for the cases that are included.

### Flagging: SCIPOP Battery

```{r}
# Focal flagging on Science Populism — the primary construct of this assignment.
# A case is flagged if: ii_sd == 0 (straightlining) OR mahal in bottom/top 10%.

p10_scipop <- quantile(rd_scipop$mahal, 0.10, na.rm = TRUE)
p90_scipop <- quantile(rd_scipop$mahal, 0.90, na.rm = TRUE)

rd_scipop_flagged <- rd_scipop |>
  mutate(
    flag_straightline = ii_sd == 0,
    flag_mahal        = (!is.na(mahal)) & (mahal <= p10_scipop | mahal >= p90_scipop),
    flag_rd           = as.integer(flag_straightline | flag_mahal)
  )

rd_scipop_flagged |>
  summarise(
    n_total         = n(),
    n_straightline  = sum(flag_straightline, na.rm = TRUE),
    n_mahal_extreme = sum(flag_mahal,        na.rm = TRUE),
    n_flagged       = sum(flag_rd,           na.rm = TRUE),
    prop_flagged    = mean(flag_rd,          na.rm = TRUE)
  )
```

The two criteria overlap substantially: most straightliners (ii_sd = 0) also fall in the bottom-10% Mahalanobis group, since a flat uniform profile is a special case of a profile close to the modal centroid. Combining the two flags therefore does not double the flagged count. The joint criterion identifies a share of US respondents whose SCIPOP response profiles give reason for concern. These cases are candidates for exclusion or sensitivity testing in subsequent analyses, where they can be cross-referenced with flags from other quality indicators (attention checks, paradata) to identify respondents who triggered multiple warning signs.



## Response Styles (Josefina Sommer)

The last kind of data quality indicators are response styles. Which style(s) do you want to focus on and how do you flag concerning cases?

To enable coherent analyses of response styles (1.) and patterns (2.), the longest possible sequence of 5pt Likert scale items entailing our focal scale - the SCIPOP scale - was used. I deliberately chose not to harmonize these items, allowing the authentic capture of natural differences in scale use across items without introducing artefactual distortions through rescaling. This preserves the ordinal nature of the data as well as item-specific variance, which is essential for the precise identification of patterns.

### 1. Response Styles
```{r}
# 1: Response styles
all5pt <- data_raw %>%
select(ID_QUALTRICS, starts_with("GOALS_PRIO_"),
starts_with("GOALS_TACKLE_"),
starts_with("NORMPERC_"),
starts_with("WILLVUL_"),
TRUST_METHOD,
TRUST_PEW,
starts_with("OUTSPOKEN_"),
starts_with("SCIPOP_"))

# Numeric IDs
all5pt <- all5pt %>%
mutate(ID_num = as.integer(row_number()))

all5pt_only <- all5pt %>%
select(starts_with("GOALS_PRIO_"),
starts_with("GOALS_TACKLE_"),
starts_with("NORMPERC_"),
starts_with("WILLVUL_"),
TRUST_METHOD,
TRUST_PEW,
starts_with("OUTSPOKEN_"),
starts_with("SCIPOP_")) %>%
mutate(across(everything(), ~ as.integer(as.numeric(.)))) # Safe Integer-Konversion

all5pt_RS <- all5pt_only %>%
  resp_styles(
    scale_min = 1,
    scale_max = 5,
    id = all5pt$ID_num
  )

# Missings?
all5pt_RS_missing <- all5pt_RS %>%
  select(-id) %>%       
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "Indicator",
               values_to = "Missing_Count") %>%
  mutate(
    Total = nrow(all5pt),                
    Missing_Percent = Missing_Count / Total * 100
  )

all5pt_RS_missing 

# rerun analyses with allowing 50% of items to be unanswered to check if there are less missings then
min_prop <- 0.5

all5pt_RS2 <- all5pt_only %>%
  resp_styles(
    scale_min = 1,
    scale_max = 5,
    id = all5pt$ID_num,
    min_valid_responses = min_prop
  )

all5pt_RS2_missing <- all5pt_RS2 %>%
  select(-id) %>%       
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "Indicator",
               values_to = "Missing_Count") %>%
  mutate(
    Total = nrow(all5pt),                
    Missing_Percent = Missing_Count / Total * 100
  )

all5pt_RS2_missing 

# does not change much (from 23.3% to 22.0% => decide to drop all of them)

```

23.3% of U.S. participants do not have response style indicators due to missing answers, as the minimum proportion of valid responses was set to 100%. Reducing this threshold (e.g., to 50%) only marginally decreased the proportion of missing response style indicators (to approximately 22%). Given this negligible gain in sample size, alongside the increased arbitrariness in threshold selection, I retained the more stringent criterion of complete responses and excluded all cases with any missing values.
Accordingly, only respondents who answered all 5-point items (GOALS_PRIO_, GOALS_TACKLE_, NORMPERC_, WILLVUL_, etc.) were included in the response style analyses, resulting in the exclusion of roughly 23% of the sample. The subsequent exploratory analyses of response styles and response patterns are therefore based on the remaining 3.562 respondents.

```{r}
all5pt_RS_clean <- all5pt_RS %>%
  drop_na()

# Summary
str(all5pt_RS_clean)
head(all5pt_RS_clean)

# Plot Response styles
all5pt_RS_clean %>%
  select(-id) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(value*100))+
  geom_histogram(fill = "steelblue", color = "white") +
  facet_wrap(vars(name))+
  labs(x = "% of respondents responses",
       y = "Count")+
  theme_classic(base_size = 15)
```

For ARS (tendency to agree), responses are spread across the full range, with moderate peaks at both low and high percentages, indicating that some respondents rarely agree uniformly, while others tend to agree frequently.

DRS (tendency to disagree) shows a strong concentration near 0%, meaning that very few respondents systematically disagree with items, while some show moderate use of this style.

ERS (extreme response style) also spans a wide range but has visible peaks toward 0% and 100%, suggesting that some respondents either avoid or predominantly use extreme endpoints.

MRS (midpoint response style) shows higher frequencies near 0% and 100%, indicating that many respondents either rarely or frequently choose the midpoint option.

NERS (non-extreme response style) similarly displays peaks toward both extremes, implying that respondents either rarely or frequently avoid extreme categories.

### 2. Response patterns
```{r}
# 2: Response patterns
all5pt_RP <- all5pt_only %>%
  resp_patterns(
    id = all5pt$ID_num
    )

str(all5pt_RP)
head(all5pt_RP)

# Missings?
 all5pt_RP %>%
     select(-id) %>%
   summarise(across(everything(), ~ sum(is.na(.)))) %>%
   pivot_longer(everything(),
               names_to = "Indicator",
                values_to = "Missing_Count") %>%
   mutate(
     Total = nrow(all5pt_RP),
     Missing_Percent = Missing_Count / Total * 100
   )
# Missing_patterns, drop
 all5pt_RP_clean <- all5pt_RP %>%
   drop_na() # see above
 
```

To flag potentially suspect respondents, I focused on two response pattern indicators: longest string length and number of transitions between different response options.
Longest string length captures over‑consistent response patterns, while the number of transitions indicate how much respondents switch between options.

``` {r}
# 3: Flagging

# Data‑driven flagging of potentially suspect respondents
all5pt_RP_clean <- all5pt_RP_clean %>%
  mutate(
    thresh_long = quantile(longest_string_length, 0.95),
    thresh_low_trans = quantile(n_transitions, 0.05),
    flag_suspect = as.numeric(
      longest_string_length >= thresh_long |
      n_transitions <= thresh_low_trans
    )
  )

thresholds <- all5pt_RP_clean %>%
  slice(1) %>%  
  select(thresh_long, thresh_low_trans)

thresholds

# Attach the flag to all5pt_RS_clean
all5pt_RS_clean <- all5pt_RS_clean %>%
  left_join(
    all5pt_RP_clean %>% select(id, flag_suspect),
    by = "id"
  )

# How many respondents are flagged as suspicious?
all5pt_RP_clean %>%
  summarise(
    n_total = n(),
    n_suspect = sum(flag_suspect),
    prop_suspect = mean(flag_suspect)
  )

# Plot Response patterns
all5pt_RP_clean %>%
  select(-id, -thresh_long, -thresh_low_trans) %>%
  pivot_longer(
    cols = c(n_transitions, longest_string_length),
    names_to = "indicator",
    values_to = "value"
  ) %>%
  mutate(
    indicator = case_when(
      indicator == "longest_string_length" ~ "Longest string",
      indicator == "n_transitions" ~ "Transitions",
      TRUE ~ indicator
    )
  ) %>%
  ggplot(aes(value, fill = factor(flag_suspect))) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(vars(indicator), scales = "free_x", nrow = 1) +
  scale_fill_manual(
    values = c("0" = "steelblue", "1" = "firebrick"),  
    labels = c("0" = "Not suspicious", "1" = "Suspicious"), 
    name = "Status" 
  ) +
  labs(x = "Indicator value", y = "Number of respondents") +
  theme_classic(base_size = 15)

```

For each indicator, the top 5% or bottom 5% of the distribution were defined as ‘extreme’.
A respondent was flagged as potentially suspect if at least one of these two indicators was in the respective extreme tail.
This approach emphasizes unusual response patterns rather than specific response styles (e.g., extreme or mid‑point responding) and may therefore be less confounded with substantive attitudes. I flagged potentially careless respondents complemented by low transition counts, capturing monotonous responding.

Respondents in the extreme 5% tails were marked as suspicious: top 5% longest string length OR bottom 5% transitions. 
This data-driven approach prioritizes response patterns over style indicators (ERS/MRS), reducing confounding with substantive attitudes.

Out of 3.562 complete cases, this identified 259 respondents (roughly 7%) as potentially careless.

## All together (ALL)

Now that you have different flags you may want to combine them. Which combinations do you deem suitable? And how to (partly) filter the identified cases?

For each case, the total number of triggered quality flags was calculated. Respondents were marked for exclusion when at least *two* indicators suggested problematic response behavior.

Our quality flags were:

+ Attention Check: failing the second attention check as defined above
+ Para Data: `speeding_factor > 2`
+ ...

```{r}
final_flags <- data_raw %>%
  mutate(
    # include all existing quality indicators
    flag_attention = attentioncheck_flag,
    flag_paradata = paradata_flag,
  ) %>%
  mutate(
    # count number of triggered flags per respondent
    n_flags = rowSums(across(starts_with("flag_")), na.rm = TRUE),

    # final exclusion rule
    final_exclude = n_flags >= 2
  )

final_flags %>%
  summarise(
    "N" = n(),
    "excluded (n)" = sum(final_exclude, na.rm = TRUE),
    "excluded [%]" = mean(final_exclude, na.rm = TRUE) * 100,
    "Flags (M)" = mean(n_flags, na.rm = TRUE)
  )
```

## Discuss (ALL)

The combined exclusion rule suggests that only respondents who triggerd at least two flags are removed from the analytical sample. This decision rule prioritizes specificity over sensitivity, thereby minimizing the risk of over-excluding valid but unusual respondents, for example since speeding alone may reflect familiarity with the questionnaire rather than careless responding, combining it with another indicator, e.g. attention-check, failure provides a stronger indication of insufficient data quality.

As a robustness check, future analyses could compare substantive results with and without flagged cases to evaluate whether the exclusion materially changes key estimates.

## References (ALL)

::: {#refs}
:::