---
title: "KODAQS Academy | Course D2 Survey"
subtitle: "*Assignment Task Group USA*"

date: today
date-format: "DD.MM.YYYY"
date-modified	: last-modified
author:
  - name: Josefina Sommer
  - name: Elnari Potgieter
  - name: David Simon
  - name: Ana Cristancho
  - name: Erik Paessler
format:
  html:
    echo: true
    warning: false
    message: false 
    code-overflow: scroll
    code-copy: true
    code-fold: false
    toc: true
    toc-location: left
    theme: cosmo
    html-math-method: katex
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
bibliography: references.bib
---


## Introduction (ALL)

*Science-related populism* has been conceptualized as a set of ideas suggesting a fundamental conflict between an allegedly virtuous people and an allegedly immoral academic elite over who should be in charge of science-related decision-making and over what is deemed true knowledge [@mede2020].

Conceptually, science-related populism consists of four major components: 

+ the ordinary people, 
+ the academic elite, 
+ science-related decision-making sovereignty, and 
+ truth-speaking sovereignty.

This is the  final assignment task of KODAQS Academy, Course D2 Survey of "Team USA" <br>

It contains sections for several kinds of data quality indicators  in regards to the construct above. 

Some parts of the assignment were intended to be completed collaboratively and others individually, each teammate was responsible for at least one section of the indicator. Overall, we worked together constructively and collaboratively on this script. Consequently, the listed names do not fully capture the collaborative nature of the work, as all team members contributed across multiple sections.

## Data Preparation (ALL)

In a first step you have to define your R setup (e.g., install und load relevant packages) and need to read in the dataset. Do your basic cleanup and wrangling here (e.g. filtering for your country group, selection of used variables etc.). 

### Load Packages and Read Data
```{r}
####  ------------- Load Packages ------------- ####
library(here)           # find files
library(tidyverse)      # data management
library(skimr)          # descriptives
library(labelled)       # data labels
library(psych)
library(stringr)
library(resquin)
library(gt)
library(ggcorrplot)

####  ------------- Read Data ------------- ####

# data_raw
data_raw <- read.csv(here("00_data/ds_full.csv"), sep = ";") %>%
  filter(COUNTRY_CODE == "USA")  %>%
  mutate(across(where(is.numeric), ~ na_if(., 99))) %>%
  mutate(across(where(is.character), ~ na_if(., "99")))

# raw data needs to be processed (a liitle bit)
# Source: https://osf.io/5c3qd/files/xz8bw

# (1.) Identify and exclude duplicate completes

# Initialize ID data (stored in separate csv file)
ds_id <- read.csv(here("00_data/ds_id.csv"), sep = ";") 

data_raw_tmp <- left_join(data_raw, ds_id, by = "ID_QUALTRICS") %>%
  group_by(ID_POLLCMPNY) %>%
  filter(is.na(ID_POLLCMPNY) | row_number(Recorded) == 1) %>%
  ungroup()

# Confirm that there are no duplicate completes (empty tibble)
data_raw_tmp %>% 
  filter(!is.na(ID_POLLCMPNY)) %>%
  group_by(ID_POLLCMPNY) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)

# % removed
per_rem1 <- ((nrow(data_raw) - nrow(data_raw_tmp)) / nrow(data_raw)) * 100

data_raw <- data_raw_tmp
rm(ds_id)
rm(data_raw_tmp)

# (2.) Exclude outlier values for age
# % outliers
per_out <- mean(data_raw$DEM_AGE < 18 | data_raw$DEM_AGE > 100, na.rm = TRUE) * 100

# Set to NA
data_raw <- data_raw %>%
  mutate(
    DEM_AGE = if_else(DEM_AGE < 18 | DEM_AGE > 100, 
                      as.numeric(NA),
                      DEM_AGE)         
  )

# (3.) Gender (1 = male)
data_raw <- data_raw  %>%
  mutate(DEM_GENDER = ifelse(DEM_GENDER == 1, 1, 0)) %>%
  mutate(DEM_GENDER = as.factor(DEM_GENDER))

# (4.) Exclude respondents who did not reach the first question = DEM_GENDER
data_raw_tmp <- data_raw %>%
  filter(!is.na(DEM_GENDER))

# % removed
per_rem2 <- ((nrow(data_raw) - nrow(data_raw_tmp)) / nrow(data_raw)) * 100

data_raw <- data_raw_tmp
rm(data_raw_tmp)

# data
data <- read.csv(here("00_data/ds_final.csv"), sep = ";") %>%
  filter(COUNTRY_CODE == "USA")
```

The following packages are used in this script:

```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

Two datasets are read:

- `data_raw`: The raw dataset before any cleaning or transformations.
- `data`: The analysis-ready dataset, including post-stratification.

Not all calculations can be performed using the pre-processed data, as some information is already missing. Therefore, the raw data were processed as follows:

1. Duplicate respondents were excluded: `r round(per_rem1, 2)`%
2. Age values outside the range of 18 to 100 were set to NA: `r round(per_out, 2)`%
3. Gender was recoded into a binary variable (1 = male), with self-described responses excluded from this coding.
4. Only participants were included who reached the beginning of the actual questionnaire (= question 1: DEM_GENDER): `r round(per_rem2, 2)`%

We performed all calculations on this dataset `data_raw`.

### Compute Indices

*Science-related populist attitudes* were measuered with 8 items on 4 dimensions on a 5-point scale each (1 = strongly disagree to 5 = strongly agree) [@mede2020], aggregated with the Goertz approach [@wuttke2020]. The script was privided by Mede et al. [-@mede2025].

```{r}
# SCIPOP_goertz: Science-related populist attitudes

# Compute overall and dimension scores

data_raw <- data_raw %>%
  mutate(
    SCIPOP_PPL_m = rowMeans(across(c(SCIPOP_common, SCIPOP_good)), na.rm = TRUE),
    SCIPOP_ELI_m = rowMeans(across(c(SCIPOP_advantage, SCIPOP_cahoots)), na.rm = TRUE),
    SCIPOP_DEC_m = rowMeans(across(c(SCIPOP_influence, SCIPOP_involved)), na.rm = TRUE),
    SCIPOP_TRU_m = rowMeans(across(c(SCIPOP_lifeexp, SCIPOP_rely)), na.rm = TRUE),
    SCIPOP_goertz = pmin(SCIPOP_PPL_m, SCIPOP_ELI_m, SCIPOP_DEC_m, SCIPOP_TRU_m, na.rm = TRUE)
  )

data_raw <- labelled::set_variable_labels(
  data_raw,
  SCIPOP_goertz = "Science-related populist attitudes: overall (Goertz score)",
  SCIPOP_PPL_m = "Science-related populist attitudes: conceptions of the ordinary people (mean)",
  SCIPOP_ELI_m = "Science-related populist attitudes: conceptions of the academic elite (mean)",
  SCIPOP_DEC_m = "Science-related populist attitudes: demands for decision-making sovereignty (mean)",
  SCIPOP_TRU_m = "Science-related populist attitudes: demands for truth-speaking sovereignty (mean)"
)
```


## Descriptive Analysis (ALL)

Provide basic descriptive statistics for your (sub)sample and instrument(s).

```{r}
data_des <- data_raw  %>%
  select(DEM_AGE, DEM_GENDER, # demographics
         SCIPOP_common, SCIPOP_good, SCIPOP_advantage, SCIPOP_cahoots, SCIPOP_influence, SCIPOP_involved, SCIPOP_lifeexp, SCIPOP_rely,# items
         SCIPOP_PPL_m, SCIPOP_ELI_m, SCIPOP_DEC_m, SCIPOP_TRU_m, SCIPOP_goertz) %>% # scales
  skim()

data_des
```


## Attention Checks (David)

The first kind of data quality indicators are check items. What check items are provided in the dataset and did participants pass the check?

```{r}
attcheck1.fails <- function(attcheck1){
  n_valid <- sum(!is.na(attcheck1), na.rm = T)
  n_fails <- sum(attcheck1 != "213", na.rm = T)
  return((1 - ((n_valid - n_fails) / n_valid)) * 100)
}


attcheck2.fails <- function(attcheck2){
  n_valid <- sum(!is.na(attcheck2), na.rm = T)
  n_fails <- sum(attcheck2 %in% 2:5, na.rm = T) + sum(attcheck2 == -99, na.rm = T)
  return((1 - ((n_valid - n_fails) / n_valid)) * 100)
}
```

Using the provided code by Mede et al. [-@mede2025], the first attention of writing “213” into a text box was failed by `r round(attcheck1.fails(data_raw$ATTCHECK_NUMBER), 0)`% of participants and

`r round(attcheck2.fails(data_raw$ATTCHECK_RES), 0)`% of participants who reached the second attention check of selecting “strongly disagree” for an extra item in a scale of science-related populist attitudes did not pass it.

So, the data set is partially filtered by design: Participants who failed the first attention check were redirected to the end of the survey and did not participate further. Consequently, they are not included among those who answered the scale of science-related populist attitudes.

The second attention check was part of the scale of science-related populist attitudes and is therefore particularly interesting for this assignment. Participants who failed the second attention check were also redirected to the end of the survey. The attention check was the last item on the scale. 

Several participants answered the questions incorrectly and were flagged as having failed by the authors. One could argue about whether these were truly careless respondents. This is particularly questionable for those who wrote the correct number in words instead of as a number, e.g., "two hundred thirteen.", one person used a lowercase L instead of 1. Others are more likely to be careless, e.g. "Very nice to have a good time at your new office and I love love you bye bye good mama bye mama good night too love love you". Others stay unclear or "missing", e.g. "i don't know". The following table provides an overview about the frequency: 

```{r}
attentioncheck_clean <- data_raw$ATTCHECK_NUMBER %>%
  str_to_lower() %>%                       # all lowercase
  str_trim() %>%                           # remove spaces
  str_replace_all("’", "'") %>%            # remove apostrophes
  str_replace_all("[[:punct:]]", " ") %>%  # remove punctuation characters
  str_squish()                             # remove extra spaces

attentioncheck_table <- case_when(
  
    # missing
  is.na(attentioncheck_clean) | attentioncheck_clean %in% c("", "none", "i don t know") | attentioncheck_clean == "99"
  ~ "Missing",
  
  # numeric variants
  str_detect(attentioncheck_clean, "\\b213\\b|\\b2l3\\b") 
  ~ "Correct",
  
  # numeric split with extra spaces
  str_detect(attentioncheck_clean, "^2\\s*1\\s*3$") 
  ~ "Correct",
  
  # word variants
  str_detect(attentioncheck_clean,
             "two.*hund.*thirt|two.*thirt|two.*one.*three|twoh.*thirt|twothirt|twoonethree|doscientos trece") 
  ~ "Correct: written in words",
  
  # incorrect
  TRUE ~ "Incorrect"
)

table <- table(attentioncheck_table)
prop <- prop.table(table(attentioncheck_table))
table
```
The number of failures on the second attention check decreases from `r round(attcheck2.fails(data_raw$ATTCHECK_RES), 0)`% to `r round(prop.table(prop)["Incorrect"] * 100, 0)`%.

These are candidates of carless respondents and therefore "flagged" in this section.

## Para Data (Elnari Potgieter)

This section looks at the response time indices of the USA data set, considering total response time, average time per item, relative completion time and speeding factor. This is a draft version.

1. Total Response Time

Total response time refers to the time it takes a respondent to complete the survey. The data set contains a "Duration" variable, analysed here, capturing response time in seconds. 

```{r}
# Key Descriptives of Duration variable (in seconds)
describe (data_raw$Duration)
```
Mean (1174.57) larger than median (905), showing right / positive skew in data. Further analysis to look out for high value outliers. Minimum at 133 seconds (2.22 minutes), and maximum at 26943 seconds (about 7 and a half hours).

The results do not appear to show any missing / incorrectly coded data, so no need for further data cleaning. Further consideration to percentiles / quartiles.

```{r}
# Consider quartiles
skim(data_raw$Duration)
```
Minimum (133) lower p25 (591), maximum (26953) much higher than mean (1174.57) and p75 (1341.25). The extreme high outliers may be due to opening the survey and only answering the survey later. This does not necessarily mean they were not paying attention when they were answering, so I don't think their answers are necessarily invalid. However, on the lower end, 2.22 minutes for a survey with 111 items (Mede et al., 2025) - some of which open ended - sounds very short. 

Box plots (in seconds) for further investigation.

```{r}
# Boxplot using base R
boxplot(data_raw$Duration,
        main = "Boxplot of Durations",
        ylab = "Duration (seconds)",
        col = "lightblue",
        border = "darkblue")
```
2. Average Time per Item

The average time per item indicator captures the average (mean) time spent on survey questions.

```{r}
# Number of items
num_items <- 111

# Calculate the average time per item for each respondent
average_time_per_item <- data_raw$Duration / num_items

# Compute the mean of all average times per item
mean_average_time_per_item <- mean(average_time_per_item)

# Output the results
cat("Average time per item for each respondent (in seconds):\n")
cat("\nMean of all average times per item (in seconds):", mean_average_time_per_item, "seconds\n")
```
The average time per item is 10.58 seconds. However, the extreme high outliers may impact this result. 

We proceed by looking at the number and proportion of respondents with an average time per item below 2 seconds (as per the course example) - this cut off can be changed for the final version. 

```{r}
# Identify respondents who fall below the cutoff of 2 seconds (as per example in course)
cutoff <- 2
below_cutoff <- which(average_time_per_item < cutoff)

# Count of respondents below the cutoff
count_below_cutoff <- length(below_cutoff)

# Proportion of respondents below the cutoff
proportion_below_cutoff <- count_below_cutoff / length(data_raw$Duration)

cat("\nRespondents below the cutoff of", cutoff, "seconds:\n")
print(below_cutoff)

cat("\nCount of respondents below the cutoff:", count_below_cutoff, "\n")
cat("Proportion of respondents below the cutoff:", proportion_below_cutoff, "\n")
```
Only 12 respondents (0.005%) had an average time per item below 2 seconds. 


Further steps: 
Decide whether to run the average time analysis without high-end outliers, and then identify those with significantly slower average times per item. 
Decide whether the cut off should be MRT - 4.4 seconds - per item (Andreadis, 2021)

3. Relative Time Completion

Relative time completion compares the speed of completing the questionnaire for a respondent to the median completion time of the sample. 

```{r}
# Median completion time of the sample
median_time <- 905

# Calculate the relative completion time for each respondent
relative_completion_time <- data_raw$Duration/ median_time

# Output the results
cat("Relative completion time for each respondent:\n")

# Summary statistics for relative completion times
cat("\nSummary of relative completion times:\n")
print(summary(relative_completion_time))
```
As per the course example, we calculate how many respondents are faster than twice the median completion time. 
Please note my findings here need to be revised. I need to check if that is median response time or median of the relative completion time. 
```{r}
# Identify respondents who are faster than two times the median completion time
cutoff <- 2
faster_than_cutoff <- which(relative_completion_time > cutoff)

# Count of respondents faster than two times the median completion time
count_faster <- length(faster_than_cutoff)

# Proportion of respondents faster than two times the median completion time
proportion_faster <- count_faster / length(data_raw$Duration)

# Identify rows where respondents are twice as fast as the median response time
twice_faster_rows <- which(relative_completion_time < 2)

# Output the results
cat("\nRespondents faster than two times the median completion time:\n")
cat("\nCount of respondents faster than two times the median:", count_faster, "\n")
cat("Proportion of respondents faster than two times the median:", proportion_faster, "\n")
cat("\nRows where respondents are twice as fast as the median response time:\n")
```
317 (13%) respondents where twice as fast as the median of the relative completion time. I did not print the rows for now, as they are quite many. 

4. Speeding Factor

Leiner, 2019 suggests a factor of larger than two as outliers.This is calculated using their response time in relation to the median of the sample. 

```{r}
# Speeding factor
# Calculate the speeding factor for each respondent
speeding_factor <- median_time / data_raw$Duration

# Output the results
cat("Speeding factor for each respondent:\n")

# Summary statistics for speeding factors
cat("\nSummary of speeding factors:\n")
print(summary(speeding_factor))
```
Some outliers with a speeding factor above 2, when considering maximum (6.8).

```{r}
# Identify outliers (speeding factor > 2)
outliers <- which(speeding_factor > 2)

# Count of outliers
outlier_count <- length(outliers)

# Proportion of outliers
outlier_proportion <- outlier_count / length(data$Duration)

# Output the count and proportion of outliers
cat("\nCount of outliers (speeding factor > 2):", outlier_count, "\n")
cat("Proportion of outliers:", outlier_proportion, "\n")
```
348 (14%) respondents with speeding factor above 2. I did not print the rows for now, as there are quite many.

5. Summary

Suggestion to compare the rows highlighted throgugh this analysis with the rows identified in other analysed (for example attention checks) to identify respondents which may need to be flagged. 

6. References

Knopf T., & Roth M. (2025). Tools and Workflows for Survey Data – Paradata [Slide deck]. In F. Bemmann, J. Blumenberg, J. Daikeler, F. Draxler, M. Fritz, F. Keusch, F. Krämer, S. Jünger, S. Kapidzic, T. Knopf, F. Kreuter, V. Kunz, L. Mejia Lopez, Y. Peters, B. Rammstedt, P. Siegers, J. Sun, C. Wagner, K. Weller, C. E. Yildiz, & L. Young (Eds.), KODAQS Academy.

Leiner, Dominik J.. “Too Fast, too Straight, too Weird: Non-Reactive Indicators for Meaningless Data in Internet Surveys.” Survey research methods 13 (2019): 229-248.

Mede, N.G., Cologna, V., Berger, S. et al. Perceptions of science, science communication, and climate change attitudes in 68 countries – the TISP dataset. Sci Data 12, 114 (2025). https://doi.org/10.1038/s41597-024-04100-7

## Missing Data (Ana)

The third kind of data quality indicators are missing data. How much data is missing, which data missing pattern occurs and are there cases with many missing data? (in respect to your choosen instrument(s) / variable(s)).

```{r}
####  ------------- Load Packages ------------- ####
library(resquin)
library(visdat)
library(here) 
library(dplyr)
library(tibble)
library(naniar)

ds_usa <- read.csv(here("00_data/ds_full.csv"), sep = ";") %>%
  filter(COUNTRY_CODE == "USA")

trust_sci_usa <- ds_usa |>
  dplyr::select(dplyr::starts_with("TRUST_SCI_"))

clim_gov_usa <- ds_usa |>
  dplyr::select(dplyr::starts_with("CLIM_GOV_"))

clim_emo_usa <- ds_usa |>
  dplyr::select(dplyr::starts_with("CLIM_EMO_"))

scipop_usa <- ds_usa |>
  dplyr::select(dplyr::starts_with("SCIPOP_"))

missing_diag_usa <- ds_usa |>
  dplyr::select(
    dplyr::starts_with("TRUST_SCI_"),
    dplyr::starts_with("CLIM_GOV_"),
    dplyr::starts_with("CLIM_EMO_"),
    dplyr::starts_with("SCIPOP_")
  )

```

```{r}
visdat::vis_miss(missing_diag_usa)

```
The missingness map indicates that nonresponse is structured rather than randomly dispersed across items. Missing values cluster in recurring horizontal bands and battery-level blocks, suggesting that missingness is driven primarily by section-level omission, routing, or broader respondent-level nonresponse rather than isolated item skipping.

```{r} 
naniar::gg_miss_var(missing_diag_usa, show_pct = TRUE)

```

The trust-in-scientists battery shows moderate but highly uniform missingness across items, with each item missing in roughly 22% of cases. This suggests that missingness in this battery is not item-specific, but instead reflects a common section-level process affecting all trust items similarly.

```{r}
naniar::gg_miss_upset(missing_diag_usa)
VIM::aggr(missing_diag_usa) |> plot(numbers = TRUE, prop = FALSE)

```

Missingness is concentrated in recurring blocks of variables rather than dispersed randomly across items. This suggests structured battery- or section-level missingness rather than simple random item-level nonresponse. 

#Person-level Missingness

```{r}
n_miss_per_case <- is.na(missing_diag_usa) |>
  as_tibble() |>
  rowSums()

missing_cases_usa <- tibble(
  respondent_id = seq_len(nrow(missing_diag_usa)),
  n_missing = n_miss_per_case
)

summary(missing_cases_usa$n_missing)

```
Respondent-level missingness across the combined batteries was highly uneven. While at least half of respondents had no missing values, the upper quartile was missing 16 or more items, and the maximum was 36 missing items.
```{r}
missing_cases_usa |>
  dplyr::filter(n_missing == ncol(missing_diag_usa)) |>
  print(n = 100)

naniar::gg_miss_case(missing_diag_usa, show_pct = TRUE)

missing_cases_usa |>
  dplyr::filter(n_missing > 0) |>
  dplyr::arrange(dplyr::desc(n_missing)) |>
  print(n = 100)

missing_cases_usa |>
  dplyr::summarise(
    n_with_any_missing = sum(n_missing > 0),
    pct_with_any_missing = mean(n_missing > 0) * 100
  )

missing_cases_usa |>
  dplyr::count(n_missing) |>
  dplyr::arrange(dplyr::desc(n_missing))

ncol(missing_diag_usa)

```


```{r}
battery_missing_summary <- tibble(
  battery = c("TRUST_SCI", "CLIM_GOV", "CLIM_EMO", "SCIPOP"),
  n_items = c(
    ncol(trust_sci_usa),
    ncol(clim_gov_usa),
    ncol(clim_emo_usa),
    ncol(scipop_usa)
  ),
  mean_pct_missing = c(
    mean(is.na(as.matrix(trust_sci_usa))) * 100,
    mean(is.na(as.matrix(clim_gov_usa))) * 100,
    mean(is.na(as.matrix(clim_emo_usa))) * 100,
    mean(is.na(as.matrix(scipop_usa))) * 100
  )
)

battery_missing_summary

```


#Item-Level Missingness

```{r}
item_missing_usa <- missing_diag_usa |>
  dplyr::summarise(
    across(everything(), ~ mean(is.na(.x)) * 100)
  ) |>
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "pct_missing"
  ) |>
  dplyr::arrange(dplyr::desc(pct_missing))

item_missing_usa

ggplot(item_missing_usa, aes(x = reorder(variable, pct_missing), y = pct_missing)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    x = "Variable",
    y = "% missing",
    title = "Item-level missingness in selected USA survey batteries"
  ) +
  theme_minimal()

```

The trust-in-scientists battery shows moderate but very even missingness across items, suggesting that nonresponse reflects a common battery-level process rather than selective skipping of particular trust items.

#Domain-level missingness

```{r}
domain_missing_usa <- ds_usa |>
  dplyr::transmute(
    miss_trust_sci = as.integer(dplyr::if_all(dplyr::starts_with("TRUST_SCI_"), is.na)),
    miss_clim_gov  = as.integer(dplyr::if_all(dplyr::starts_with("CLIM_GOV_"), is.na)),
    miss_clim_emo  = as.integer(dplyr::if_all(dplyr::starts_with("CLIM_EMO_"), is.na)),
    miss_scipop    = as.integer(dplyr::if_all(dplyr::starts_with("SCIPOP_"), is.na))
  )

domain_missing_usa |>
  dplyr::summarise(dplyr::across(everything(), sum))

cor(domain_missing_usa)
round(cor(domain_missing_usa), 3)


```
Battery-level missingness indicators were very strongly correlated, especially between climate-government and climate-emotion items (r = 0.997) and between trust-in-scientists and science-populism items (r = 0.978). This indicates that missingness is driven by structured module-level processes rather than isolated or random nonresponse.

#Complete-case Restriction
```{r}
complete_cases_usa <- missing_diag_usa |>
  tidyr::drop_na()

nrow(missing_diag_usa)
nrow(complete_cases_usa)

tibble(
  total_cases = nrow(missing_diag_usa),
  complete_cases = nrow(complete_cases_usa),
  cases_dropped = nrow(missing_diag_usa) - nrow(complete_cases_usa),
  pct_dropped = (nrow(missing_diag_usa) - nrow(complete_cases_usa)) / nrow(missing_diag_usa) * 100
)

```
Complete-case analysis would exclude 2,067 of 4,662 respondents (44.3%), showing that listwise deletion is highly costly when these batteries are analysed together, , indicating that complete-case analysis would substantially reduce statistical power and is poorly suited to a dataset in which missingness is structured at the battery level.

#Logistic Regression
```{r}

# install.packages("logistf")
library(logistf)


miss_model_trust_sci <- glm(
  miss_trust_sci ~ DEM_AGE + DEM_EDU + Duration + Response_TYPE 
  + DEM_GENDER,
  data = dplyr::bind_cols(ds_usa, domain_missing_usa),
  family = binomial()
)

summary(miss_model_trust_sci)

model_df <- dplyr::bind_cols(ds_usa, domain_missing_usa) |>
  dplyr::filter(
    !is.na(DEM_AGE),
    !is.na(DEM_EDU),
    !is.na(Duration),
    !is.na(Response_TYPE),
    !is.na(DEM_GENDER)
  ) |>
  dplyr::mutate(
    DEM_EDU       = factor(DEM_EDU),
    Response_TYPE = factor(Response_TYPE),
    DEM_GENDER    = factor(DEM_GENDER)
  )

table(model_df$miss_trust_sci, model_df$Response_TYPE)
table(model_df$miss_trust_sci, model_df$DEM_GENDER)
table(model_df$miss_trust_sci, model_df$DEM_EDU)

miss_model_trust_sci <- logistf::logistf(
  miss_trust_sci ~ DEM_AGE + DEM_EDU + Duration + Response_TYPE + DEM_GENDER,
  data = model_df
)
 
summary(miss_model_trust_sci)

tibble::tibble(
  term      = miss_model_trust_sci$terms,
  odds_ratio = exp(coef(miss_model_trust_sci)),
  ci_lower  = exp(miss_model_trust_sci$ci.lower),
  ci_upper  = exp(miss_model_trust_sci$ci.upper),
  p_value   = miss_model_trust_sci$prob
)

```
A penalised logistic regression model was used to account for sparse categories and separation issues. Results indicate that missingness in the trust-in-scientists battery is systematically structured. Older respondents were significantly less likely to exhibit whole-battery missingness, and gender differences were also observed, with one group showing a lower probability of omission. Completion duration was positively associated with missingness, although the effect size was small. In contrast, education was not a robust predictor once modelled categorically. A survey-administration variable (Response Type) exhibited near-perfect separation and was therefore interpreted as a structural artefact rather than a substantive determinant of missingness.

Taken together, these findings suggest that missingness is neither Missing Completely at Random (MCAR) nor purely item-level. Instead, it reflects a combination of respondent characteristics, survey engagement, and design-driven processes operating at the battery level.

## Response Distributions (Erik Paessler)

The fourth kind of data quality indicators are response distributions. Which indicators do you deem relevant and which criterion do you use for flagging?

For response distributions, I focus on **intra-individual (within-person) location and spread indicators** computed with `resquin::resp_distributions()`. Five Likert-scale batteries are examined separately so the statistics remain interpretable within each coherent scale: Trust in Scientists (`TRUST_SCI_`), Climate Government Trust (`CLIM_GOV_`), Climate Emotions (`CLIM_EMO_`), Science Populism (`SCIPOP_`, 8 original items), and Science Information Sources (`SCIINFO_`).

The indicators I deem most relevant are:

- **`ii_sd`** (intra-individual standard deviation): captures within-person response spread. A value of zero flags **straightliners** — respondents who chose the identical option for every item in a battery, a strong marker of inattentive responding.
- **`mahal`** (Mahalanobis distance): captures how unusual a respondent's response profile is relative to the sample distribution. Respondents in the **bottom 10%** have profiles suspiciously close to the sample centroid (overly "average", consistent with random or careless uniform clicking); respondents in the **top 10%** show unusually deviant profiles.
- **`ii_mean` / `ii_median`**: secondary indicators used to characterise the nature of flagged profiles (e.g., persistent floor or ceiling responding combined with near-zero SD).

**Flagging criterion**: A case is flagged as a potential data quality concern if it meets at least one of the following for the focal SCIPOP battery: (1) `ii_sd == 0` (straightlining), or (2) Mahalanobis distance in the bottom or top 10% of the sample. All other batteries serve as comparative context.

```{r}
# Extract item batteries from data_raw ----

# Trust in Scientists
trust_sci <- data_raw |> select(starts_with("TRUST_SCI_"))

# Climate Government Trust
clim_gov  <- data_raw |> select(starts_with("CLIM_GOV_"))

# Climate Emotions
clim_emo  <- data_raw |> select(starts_with("CLIM_EMO_"))

# Science Populism: original 8 items only (exclude computed indices added earlier)
scipop <- data_raw |>
  select(SCIPOP_common, SCIPOP_good, SCIPOP_advantage, SCIPOP_cahoots,
         SCIPOP_influence, SCIPOP_involved, SCIPOP_lifeexp, SCIPOP_rely)

# Science Information Sources
sciinfo <- data_raw |> select(starts_with("SCIINFO_"))


# Compute response distribution indicators ----

rd_trust_sci <- resp_distributions(trust_sci)
rd_clim_gov  <- resp_distributions(clim_gov)
rd_clim_emo  <- resp_distributions(clim_emo)
rd_scipop    <- resp_distributions(scipop)
rd_sciinfo   <- resp_distributions(sciinfo)
```

### Descriptive Summaries

```{r}
cat(" Trust in Scientists \n");  summary(rd_trust_sci)
cat(" Climate Gov. Trust \n");   summary(rd_clim_gov)
cat(" Climate Emotions \n");     summary(rd_clim_emo)
cat(" Science Populism \n");     summary(rd_scipop)
cat(" Science Info Sources \n"); summary(rd_sciinfo)
```

The `ii_mean` values differ noticeably across batteries. Trust in Scientists and Science Information Sources tend toward the upper end of the scale, while Science Populism responses cluster more around the midpoint. Climate Emotions stands out with higher missingness, reflected in the `na_n` column — respondents more frequently skipped individual items compared to the other batteries.

### Univariate Distribution Plots

```{r}
plot_univariate <- function(rd, title) {
  rd |>
    select(-contains("na")) |>
    pivot_longer(-id) |>
    ggplot(aes(value)) +
    geom_histogram(bins = 30, fill = "steelblue", color = "white") +
    facet_wrap(vars(name), scales = "free_x") +
    labs(title = title, subtitle = "US respondents", x = NULL, y = "Count") +
    theme_minimal(base_size = 12)
}

plot_univariate(rd_trust_sci, "Response Distributions – Trust in Scientists")
plot_univariate(rd_clim_gov,  "Response Distributions – Climate Gov. Trust")
plot_univariate(rd_clim_emo,  "Response Distributions – Climate Emotions")
plot_univariate(rd_scipop,    "Response Distributions – Science Populism")
plot_univariate(rd_sciinfo,   "Response Distributions – Science Info Sources")
```

The `ii_sd` histograms show a spike at zero across all batteries — straightliners — but the spike is most pronounced for CLIM_GOV and SCIPOP. The `ii_mean` distributions differ in shape: Trust in Scientists is left-skewed (most respondents express moderate-to-high trust), whereas SCIPOP is roughly symmetric around the midpoint. Mahalanobis distances are right-skewed in all batteries, as expected under a chi-squared distribution.

### Bivariate Scatter Plots (Indicator Correlations)

```{r}
# Pairs plots showing correlations between indicators within each battery, (NA-count columns excluded)

rd_trust_sci |> select(-contains("na")) |> plot(main = "Trust in Scientists – USA")
rd_clim_gov  |> select(-contains("na")) |> plot(main = "Climate Gov. Trust – USA")
rd_clim_emo  |> select(-contains("na")) |> plot(main = "Climate Emotions – USA")
rd_scipop    |> select(-contains("na")) |> plot(main = "Science Populism – USA")
rd_sciinfo   |> select(-contains("na")) |> plot(main = "Science Info Sources – USA")
```

Across all batteries, `ii_mean` and `ii_median` are highly correlated (near-diagonal), confirming that both location indicators capture the same dimension. The relationship between `ii_sd` and `mahal` is positive but non-linear: straightliners (ii_sd ≈ 0) tend to have low Mahalanobis distances, since a flat all-same response profile often coincides with the modal response category. This pattern is clearest in the SCIPOP and CLIM_GOV batteries.

### Straightlining: All Batteries

```{r}
# Count respondents with ii_sd == 0 (indicating identical response to every item) in each battery

straight_summary <- tibble(
  Battery    = c("Trust in Scientists", "Climate Gov. Trust", "Climate Emotions",
                 "Science Populism",    "Science Info Sources"),
  n_total    = c(nrow(rd_trust_sci), nrow(rd_clim_gov), nrow(rd_clim_emo),
                 nrow(rd_scipop),    nrow(rd_sciinfo)),
  n_straight = c(
    sum(rd_trust_sci$ii_sd == 0, na.rm = TRUE),
    sum(rd_clim_gov$ii_sd  == 0, na.rm = TRUE),
    sum(rd_clim_emo$ii_sd  == 0, na.rm = TRUE),
    sum(rd_scipop$ii_sd    == 0, na.rm = TRUE),
    sum(rd_sciinfo$ii_sd   == 0, na.rm = TRUE)
  )
) |>
  mutate(pct_straight = round(n_straight / n_total * 100, 1))

print(straight_summary)
```

Straightlining rates vary across batteries. Climate Government Trust and Science Populism tend to show the highest proportions, which may reflect the ideological or evaluative nature of those items — some respondents may reject the entire construct rather than differentiate between items. Trust in Scientists and Science Information Sources generally show lower straightlining, possibly because respondents find it easier to discriminate between items with more concrete referents.

### Mahalanobis Distance: All Batteries

```{r}
mahal_summary <- function(rd, label) {
  p10 <- quantile(rd$mahal, 0.10, na.rm = TRUE)
  p90 <- quantile(rd$mahal, 0.90, na.rm = TRUE)
  cat(sprintf("\n %s \n", label))
  cat(sprintf("  10th pct cutoff: %.3f   |   90th pct cutoff: %.3f\n", p10, p90))
  cat("  Bottom 10% (uniform profiles):\n")
  print(summary(rd |> filter(mahal <= p10)))
  cat("  Top 10% (deviant profiles):\n")
  print(summary(rd |> filter(mahal >= p90)))
}

mahal_summary(rd_trust_sci, "Trust in Scientists")
mahal_summary(rd_clim_gov,  "Climate Gov. Trust")
mahal_summary(rd_clim_emo,  "Climate Emotions")
mahal_summary(rd_scipop,    "Science Populism")
mahal_summary(rd_sciinfo,   "Science Info Sources")
```

The bottom-10% Mahalanobis groups (most uniform profiles) are characterised by low `ii_sd` and `ii_mean` / `ii_median` values close to the sample mean — consistent with undifferentiated, midpoint-heavy responding. The top-10% groups show higher `ii_sd` and more extreme `ii_mean` values, indicating strong scale use in one direction. Across batteries, the SCIPOP bottom-10% group is particularly notable: the combination of low spread and near-midpoint means is consistent with careless, non-discriminating responses on a construct that requires genuine attitude engagement.

### Sensitivity Check: `min_valid_responses` Threshold

```{r}
# CLIM_EMO is used here because it contains notable item-level missings,
# making it the best illustration of the threshold trade-off

cat("CLIM_EMO: default (min_valid = 1.0)\n")
clim_emo |> resp_distributions(min_valid_responses = 1.0) |> summary()

cat("CLIM_EMO: min_valid = 0.9\n")
clim_emo |> resp_distributions(min_valid_responses = 0.9) |> summary()

cat("CLIM_EMO: min_valid = 0.5\n")
clim_emo |> resp_distributions(min_valid_responses = 0.5) |> summary()

# ii_mean and ii_median remain stable across thresholds and ii_sd and mahal become less reliable as more missings are tolerated.
```

Lowering the threshold recovers additional cases but at the cost of reliability: `ii_sd` and `mahal` become noisier when computed on incomplete item sets, as the spread and distance metrics depend on the full response vector. Since `ii_mean` and `ii_median` are robust regardless, they remain useful for partially answered batteries, but the flagging criteria based on `ii_sd` and `mahal` are applied only where all items are present.

### Flagging: SCIPOP Battery

```{r}
# Focal flagging on Science Populism
# A case is flagged if ii_sd == 0 (straightlining) OR mahal in bottom/top 10%.

p10_scipop <- quantile(rd_scipop$mahal, 0.10, na.rm = TRUE)
p90_scipop <- quantile(rd_scipop$mahal, 0.90, na.rm = TRUE)

rd_scipop_flagged <- rd_scipop |>
  mutate(
    flag_straightline = ii_sd == 0,
    flag_mahal        = (!is.na(mahal)) & (mahal <= p10_scipop | mahal >= p90_scipop),
    flag_rd           = as.integer(flag_straightline | flag_mahal)
  )

rd_scipop_flagged |>
  summarise(
    n_total         = n(),
    n_straightline  = sum(flag_straightline, na.rm = TRUE),
    n_mahal_extreme = sum(flag_mahal,        na.rm = TRUE),
    n_flagged       = sum(flag_rd,           na.rm = TRUE),
    prop_flagged    = mean(flag_rd,          na.rm = TRUE)
  )
```

The two criteria partially overlap: most straightliners also fall in the bottom-10% Mahalanobis group, so combining them does not simply double the flagged count. The combined flag identifies a small but non-trivial share of US respondents whose SCIPOP response profiles give reason for concern. These cases are candidates for removal or sensitivity analysis in subsequent steps, where they can be compared against flags from other quality indicators (attention checks, paradata) to identify respondents who triggered multiple warning signs.



## Response Styles (Josefina Sommer)

The last kind of data quality indicators are response styles. Which style(s) do you want to focus on and how do you flag concerning cases?

To enable coherent analyses of response styles (1.) and patterns (2.), the longest possible sequence of 5pt Likert scale items entailing our focal scale - the SCIPOP scale - was used. I deliberately chose not to harmonize these items, allowing the authentic capture of natural differences in scale use across items without introducing artefactual distortions through rescaling. This preserves the ordinal nature of the data as well as item-specific variance, which is essential for the precise identification of patterns.

### 1. Response Styles
```{r}
# 1: Response styles
all5pt <- data_raw %>%
select(ID_QUALTRICS, starts_with("GOALS_PRIO_"),
starts_with("GOALS_TACKLE_"),
starts_with("NORMPERC_"),
starts_with("WILLVUL_"),
TRUST_METHOD,
TRUST_PEW,
starts_with("OUTSPOKEN_"),
starts_with("SCIPOP_"))

# Numeric IDs
all5pt <- all5pt %>%
mutate(ID_num = as.integer(row_number()))

all5pt_only <- all5pt %>%
select(starts_with("GOALS_PRIO_"),
starts_with("GOALS_TACKLE_"),
starts_with("NORMPERC_"),
starts_with("WILLVUL_"),
TRUST_METHOD,
TRUST_PEW,
starts_with("OUTSPOKEN_"),
starts_with("SCIPOP_")) %>%
mutate(across(everything(), ~ as.integer(as.numeric(.)))) # Safe Integer-Konversion

all5pt_RS <- all5pt_only %>%
  resp_styles(
    scale_min = 1,
    scale_max = 5,
    id = all5pt$ID_num
  )

# Missings?
all5pt_RS_missing <- all5pt_RS %>%
  select(-id) %>%       
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "Indicator",
               values_to = "Missing_Count") %>%
  mutate(
    Total = nrow(all5pt),                
    Missing_Percent = Missing_Count / Total * 100
  )

all5pt_RS_missing 

# rerun analyses with allowing 50% of items to be unanswered to check if there are less missings then
min_prop <- 0.5

all5pt_RS2 <- all5pt_only %>%
  resp_styles(
    scale_min = 1,
    scale_max = 5,
    id = all5pt$ID_num,
    min_valid_responses = min_prop
  )

all5pt_RS2_missing <- all5pt_RS2 %>%
  select(-id) %>%       
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "Indicator",
               values_to = "Missing_Count") %>%
  mutate(
    Total = nrow(all5pt),                
    Missing_Percent = Missing_Count / Total * 100
  )

all5pt_RS2_missing 

# does not change much (from 23.3% to 22.0% => decide to drop all of them)

```

23.3% of U.S. participants do not have response style indicators due to missing answers, as the minimum proportion of valid responses was set to 100%. Reducing this threshold (e.g., to 50%) only marginally decreased the proportion of missing response style indicators (to approximately 22%). Given this negligible gain in sample size, alongside the increased arbitrariness in threshold selection, I retained the more stringent criterion of complete responses and excluded all cases with any missing values.
Accordingly, only respondents who answered all 5-point items (GOALS_PRIO_, GOALS_TACKLE_, NORMPERC_, WILLVUL_, etc.) were included in the response style analyses, resulting in the exclusion of roughly 23% of the sample. The subsequent exploratory analyses of response styles and response patterns are therefore based on the remaining 3.562 respondents.

```{r}
all5pt_RS_clean <- all5pt_RS %>%
  drop_na()

# Summary
str(all5pt_RS_clean)
head(all5pt_RS_clean)

# Plot Response styles
all5pt_RS_clean %>%
  select(-id) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(value*100))+
  geom_histogram(fill = "steelblue", color = "white") +
  facet_wrap(vars(name))+
  labs(x = "% of respondents responses",
       y = "Count")+
  theme_classic(base_size = 15)
```

For ARS (tendency to agree), responses are spread across the full range, with moderate peaks at both low and high percentages, indicating that some respondents rarely agree uniformly, while others tend to agree frequently.

DRS (tendency to disagree) shows a strong concentration near 0%, meaning that very few respondents systematically disagree with items, while some show moderate use of this style.

ERS (extreme response style) also spans a wide range but has visible peaks toward 0% and 100%, suggesting that some respondents either avoid or predominantly use extreme endpoints.

MRS (midpoint response style) shows higher frequencies near 0% and 100%, indicating that many respondents either rarely or frequently choose the midpoint option.

NERS (non-extreme response style) similarly displays peaks toward both extremes, implying that respondents either rarely or frequently avoid extreme categories.

### 2. Response patterns
```{r}
# 2: Response patterns
all5pt_RP <- all5pt_only %>%
  resp_patterns(
    id = all5pt$ID_num
    )

str(all5pt_RP)
head(all5pt_RP)

# Missings?
 all5pt_RP %>%
     select(-id) %>%
   summarise(across(everything(), ~ sum(is.na(.)))) %>%
   pivot_longer(everything(),
               names_to = "Indicator",
                values_to = "Missing_Count") %>%
   mutate(
     Total = nrow(all5pt_RP),
     Missing_Percent = Missing_Count / Total * 100
   )
# Missing_patterns, drop
 all5pt_RP_clean <- all5pt_RP %>%
   drop_na() # see above
 
```

To flag potentially suspect respondents, I focused on two response pattern indicators: longest string length and number of transitions between different response options.
Longest string length captures over‑consistent response patterns, while the number of transitions indicate how much respondents switch between options.

``` {r}
# 3: Flagging

# Data‑driven flagging of potentially suspect respondents
all5pt_RP_clean <- all5pt_RP_clean %>%
  mutate(
    thresh_long = quantile(longest_string_length, 0.95),
    thresh_low_trans = quantile(n_transitions, 0.05),
    flag_suspect = as.numeric(
      longest_string_length >= thresh_long |
      n_transitions <= thresh_low_trans
    )
  )

thresholds <- all5pt_RP_clean %>%
  slice(1) %>%  
  select(thresh_long, thresh_low_trans)

thresholds

# Attach the flag to all5pt_RS_clean
all5pt_RS_clean <- all5pt_RS_clean %>%
  left_join(
    all5pt_RP_clean %>% select(id, flag_suspect),
    by = "id"
  )

# How many respondents are flagged as suspicious?
all5pt_RP_clean %>%
  summarise(
    n_total = n(),
    n_suspect = sum(flag_suspect),
    prop_suspect = mean(flag_suspect)
  )

# Plot Response patterns
all5pt_RP_clean %>%
  select(-id, -thresh_long, -thresh_low_trans) %>%
  pivot_longer(
    cols = c(n_transitions, longest_string_length),
    names_to = "indicator",
    values_to = "value"
  ) %>%
  mutate(
    indicator = case_when(
      indicator == "longest_string_length" ~ "Longest string",
      indicator == "n_transitions" ~ "Transitions",
      TRUE ~ indicator
    )
  ) %>%
  ggplot(aes(value, fill = factor(flag_suspect))) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(vars(indicator), scales = "free_x", nrow = 1) +
  scale_fill_manual(
    values = c("0" = "steelblue", "1" = "firebrick"),  
    labels = c("0" = "Not suspicious", "1" = "Suspicious"), 
    name = "Status" 
  ) +
  labs(x = "Indicator value", y = "Number of respondents") +
  theme_classic(base_size = 15)

```

For each indicator, the top 5% or bottom 5% of the distribution were defined as ‘extreme’.
A respondent was flagged as potentially suspect if at least one of these two indicators was in the respective extreme tail.
This approach emphasizes unusual response patterns rather than specific response styles (e.g., extreme or mid‑point responding) and may therefore be less confounded with substantive attitudes. I flagged potentially careless respondents complemented by low transition counts, capturing monotonous responding.

Respondents in the extreme 5% tails were marked as suspicious: top 5% longest string length OR bottom 5% transitions. 
This data-driven approach prioritizes response patterns over style indicators (ERS/MRS), reducing confounding with substantive attitudes.

Out of 3.562 complete cases, this identified 259 respondents (roughly 7%) as potentially careless.


## All together (ALL)

Now that you have different flags you may want to combine them. Which combinations do you deem suitable? And how to (partly) filter the identified cases?

```{r}



```

## Discuss (ALL)

Discuss central findings shortly! Make some statements on your final decision in respect to include/exclude rules (cutoffs/thresholds/filter combination), data loss, consequences for planned analyses, and any notable what you like to adress.




### Response Styles and Patterns
Response styles and pattern analyses based on a broad set of 5-point Likert items revealed substantial heterogeneity in scale use across respondents. To ensure reliable estimation, analyses were restricted to complete cases, resulting in the exclusion of approximately 23% of the sample; lowering the minimum response threshold did not meaningfully reduce data loss.

While response style indicators (e.g., agreement, midpoint, and extreme responding) showed considerable variation, potentially suspect cases were primarily identified using response pattern indicators. Specifically, respondents were flagged based on extreme values in longest response strings and low numbers of response transitions, capturing overly consistent or monotonous answering behavior.

Using this conservative, data-driven approach, approximately 7% of the 3.562 respondents were identified as potentially careless. By focusing on response patterns rather than style indicators, this strategy reduces confounding with substantive attitudes and complements other data quality measures such as attention checks and paradata.


## References (ALL)

::: {#refs}
:::

::: {.callout-important appearance="default"}

## Pay Attention Please!
For final submission upload your prepared Quarto document and an exported html-file 
into Moodle, week 8, section "Final Assignment Course D2 Survey LU8.1".
**FINAL DEADLINE will be June 20th, 2025, sharp!**

:::

<span style="color:#1E8CC8;">**Good luck and see you next Thursday (June 5th) on our last weekly touchpoint!**</span>👋



