---
title: "KODAQS Academy | Course D2 Survey"
subtitle: "*Assignment Task Group USA*"

date: today
date-format: "DD.MM.YYYY"
date-modified	: last-modified
author:
  - name: Josefina Sommer
  - name: Elnari Potgieter
  - name: David Simon
  - name: Ana Cristancho
  - name: Erik Paessler
format:
  html:
    echo: true
    warning: false
    message: false 
    code-overflow: scroll
    code-copy: true
    code-fold: false
    toc: true
    toc-location: left
    theme: cosmo
    html-math-method: katex
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
bibliography: references.bib
---


## Introduction (ALL)

Other suggested intro:
This assignment explores data quality aspects of United States of America (USA) data set of the Trust in Science and Science-related Populism (TISP) project. The TISP project consists of a consortium of 241 researchers from 179 institutions focused on global public opinion regarding science. Data was collected from 68 countries from November 2022 to August 2023, with the survey gathering the opinions of 71,922 respondents in total (TISP, n.d).

The project measures various constructs and characteristics. We focus on the questionnaire items related to science-related populism for this assignment. This construct is understood as considering conflicting ideas regarding decision-making powers pertaining to science and what is considered “true knowledge” [@mede2020]. 

As our interest here is in the data quality aspects of the dataset, we consider the findings of various data quality tests - including attention checks, para data, missing data, response distribution and response styles – to discern the overall quality of the USA dataset and to highlight any respondents that should be flagged due to their response behaviours. 


*Science-related populism* has been conceptualized as a set of ideas suggesting a fundamental conflict between an allegedly virtuous people and an allegedly immoral academic elite over who should be in charge of science-related decision-making and over what is deemed true knowledge [@mede2020].

Conceptually, science-related populism consists of four major components: 

+ the ordinary people, 
+ the academic elite, 
+ science-related decision-making sovereignty, and 
+ truth-speaking sovereignty.

This is the  final assignment task of KODAQS Academy, Course D2 Survey of "Team USA" <br>

It contains sections for several kinds of data quality indicators  in regards to the construct above. 

Some parts of the assignment were intended to be completed collaboratively and others individually, each teammate was responsible for at least one section of the indicator. Overall, we worked together constructively and collaboratively on this script. Consequently, the listed names do not fully capture the collaborative nature of the work, as all team members contributed across multiple sections.

## Data Preparation (ALL)

In a first step you have to define your R setup (e.g., install und load relevant packages) and need to read in the dataset. Do your basic cleanup and wrangling here (e.g. filtering for your country group, selection of used variables etc.). 

### Load Packages and Read Data
```{r}
####  ------------- Load Packages ------------- ####
library(here)           # find files
library(tidyverse)      # data management
library(skimr)          # descriptives
library(labelled)       # data labels
library(psych)
library(stringr)

####  ------------- Read Data ------------- ####

# data_raw
data_raw <- read.csv(here("00_data/ds_full.csv"), sep = ";") %>%
  filter(COUNTRY_CODE == "USA")  %>%
  mutate(across(where(is.numeric), ~ na_if(., 99))) %>%
  mutate(across(where(is.character), ~ na_if(., "99")))

# raw data needs to be processed (a liitle bit)
# Source: https://osf.io/5c3qd/files/xz8bw

# (1.) Identify and exclude duplicate completes

# Initialize ID data (stored in separate csv file)
ds_id <- read.csv(here("00_data/ds_id.csv"), sep = ";") 

data_raw_tmp <- left_join(data_raw, ds_id, by = "ID_QUALTRICS") %>%
  group_by(ID_POLLCMPNY) %>%
  filter(is.na(ID_POLLCMPNY) | row_number(Recorded) == 1) %>%
  ungroup()

# Confirm that there are no duplicate completes (empty tibble)
data_raw_tmp %>% 
  filter(!is.na(ID_POLLCMPNY)) %>%
  group_by(ID_POLLCMPNY) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)

# % removed
per_rem1 <- ((nrow(data_raw) - nrow(data_raw_tmp)) / nrow(data_raw)) * 100

data_raw <- data_raw_tmp
rm(ds_id)
rm(data_raw_tmp)

# (2.) Exclude outlier values for age
# % outliers
per_out <- mean(data_raw$DEM_AGE < 18 | data_raw$DEM_AGE > 100, na.rm = TRUE) * 100

# Set to NA
data_raw <- data_raw %>%
  mutate(
    DEM_AGE = if_else(DEM_AGE < 18 | DEM_AGE > 100, 
                      as.numeric(NA),
                      DEM_AGE)         
  )

# (3.) Gender (1 = male)
data_raw <- data_raw  %>%
  mutate(DEM_GENDER = ifelse(DEM_GENDER == 1, 1, 0)) %>%
  mutate(DEM_GENDER = as.factor(DEM_GENDER))

# (4.) Exclude respondents who did not reach the first question = DEM_GENDER
data_raw_tmp <- data_raw %>%
  filter(!is.na(DEM_GENDER))

# % removed
per_rem2 <- ((nrow(data_raw) - nrow(data_raw_tmp)) / nrow(data_raw)) * 100

data_raw <- data_raw_tmp
rm(data_raw_tmp)

# data
data <- read.csv(here("00_data/ds_final.csv"), sep = ";") %>%
  filter(COUNTRY_CODE == "USA")
```

The following packages are used in this script:

```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

Two datasets are read:

- `data_raw`: The raw dataset before any cleaning or transformations.
- `data`: The analysis-ready dataset, including post-stratification.

Not all calculations can be performed using the pre-processed data, as some information is already missing. Therefore, the raw data were processed as follows:

1. Duplicate respondents were excluded: `r round(per_rem1, 2)`%
2. Age values outside the range of 18 to 100 were set to NA: `r round(per_out, 2)`%
3. Gender was recoded into a binary variable (1 = male), with self-described responses excluded from this coding.
4. Only participants were included who reached the beginning of the actual questionnaire (= question 1: DEM_GENDER): `r round(per_rem2, 2)`%

We performed all calculations on this dataset `data_raw`.

### Compute Indices

*Science-related populist attitudes* were measuered with 8 items on 4 dimensions on a 5-point scale each (1 = strongly disagree to 5 = strongly agree) [@mede2020], aggregated with the Goertz approach [@wuttke2020]. The script was privided by Mede et al. [-@mede2025].

```{r}
# SCIPOP_goertz: Science-related populist attitudes

# Compute overall and dimension scores

data_raw <- data_raw %>%
  mutate(
    SCIPOP_PPL_m = rowMeans(across(c(SCIPOP_common, SCIPOP_good)), na.rm = TRUE),
    SCIPOP_ELI_m = rowMeans(across(c(SCIPOP_advantage, SCIPOP_cahoots)), na.rm = TRUE),
    SCIPOP_DEC_m = rowMeans(across(c(SCIPOP_influence, SCIPOP_involved)), na.rm = TRUE),
    SCIPOP_TRU_m = rowMeans(across(c(SCIPOP_lifeexp, SCIPOP_rely)), na.rm = TRUE),
    SCIPOP_goertz = pmin(SCIPOP_PPL_m, SCIPOP_ELI_m, SCIPOP_DEC_m, SCIPOP_TRU_m, na.rm = TRUE)
  )

data_raw <- labelled::set_variable_labels(
  data_raw,
  SCIPOP_goertz = "Science-related populist attitudes: overall (Goertz score)",
  SCIPOP_PPL_m = "Science-related populist attitudes: conceptions of the ordinary people (mean)",
  SCIPOP_ELI_m = "Science-related populist attitudes: conceptions of the academic elite (mean)",
  SCIPOP_DEC_m = "Science-related populist attitudes: demands for decision-making sovereignty (mean)",
  SCIPOP_TRU_m = "Science-related populist attitudes: demands for truth-speaking sovereignty (mean)"
)
```


## Descriptive Analysis (ALL)

Provide basic descriptive statistics for your (sub)sample and instrument(s).

```{r}
data_des <- data_raw  %>%
  select(DEM_AGE, DEM_GENDER, # demographics
         SCIPOP_common, SCIPOP_good, SCIPOP_advantage, SCIPOP_cahoots, SCIPOP_influence, SCIPOP_involved, SCIPOP_lifeexp, SCIPOP_rely,# items
         SCIPOP_PPL_m, SCIPOP_ELI_m, SCIPOP_DEC_m, SCIPOP_TRU_m, SCIPOP_goertz) %>% # scales
  skim()

data_des
```


## Attention Checks (David)

The first kind of data quality indicators are check items. What check items are provided in the dataset and did participants pass the check?

```{r}
attcheck1.fails <- function(attcheck1){
  n_valid <- sum(!is.na(attcheck1), na.rm = T)
  n_fails <- sum(attcheck1 != "213", na.rm = T)
  return((1 - ((n_valid - n_fails) / n_valid)) * 100)
}


attcheck2.fails <- function(attcheck2){
  n_valid <- sum(!is.na(attcheck2), na.rm = T)
  n_fails <- sum(attcheck2 %in% 2:5, na.rm = T) + sum(attcheck2 == -99, na.rm = T)
  return((1 - ((n_valid - n_fails) / n_valid)) * 100)
}
```

Using the provided code by Mede et al. [-@mede2025], the first attention of writing “213” into a text box was failed by `r round(attcheck1.fails(data_raw$ATTCHECK_NUMBER), 0)`% of participants and

`r round(attcheck2.fails(data_raw$ATTCHECK_RES), 0)`% of participants who reached the second attention check of selecting “strongly disagree” for an extra item in a scale of science-related populist attitudes did not pass it.

So, the data set is partially filtered by design: Participants who failed the first attention check were redirected to the end of the survey and did not participate further. Consequently, they are not included among those who answered the scale of science-related populist attitudes.

The second attention check was part of the scale of science-related populist attitudes and is therefore particularly interesting for this assignment. Participants who failed the second attention check were also redirected to the end of the survey. The attention check was the last item on the scale. 

Several participants answered the questions incorrectly and were flagged as having failed by the authors. One could argue about whether these were truly careless respondents. This is particularly questionable for those who wrote the correct number in words instead of as a number, e.g., "two hundred thirteen.", one person used a lowercase L instead of 1. Others are more likely to be careless, e.g. "Very nice to have a good time at your new office and I love love you bye bye good mama bye mama good night too love love you". Others stay unclear or "missing", e.g. "i don't know". The following table provides an overview about the frequency: 

```{r}
attentioncheck_clean <- data_raw$ATTCHECK_NUMBER %>%
  str_to_lower() %>%                       # all lowercase
  str_trim() %>%                           # remove spaces
  str_replace_all("’", "'") %>%            # remove apostrophes
  str_replace_all("[[:punct:]]", " ") %>%  # remove punctuation characters
  str_squish()                             # remove extra spaces

attentioncheck_table <- case_when(
  
  # numeric variants
  str_detect(attentioncheck_clean, "\\b213\\b|\\b2l3\\b") 
  ~ "Correct",
  
  # numeric split with extra spaces / periods / commas
  str_detect(attentioncheck_clean, "^2\\s*1\\s*3$") 
  ~ "Correct",
  
  # word variants
  str_detect(attentioncheck_clean,
             "two.*hund.*thirt|two.*thirt|two.*one.*three|twoh.*thirt|twothirt|twoonethree|doscientos trece") 
  ~ "Correct: written in words",
  
  # missing
  attentioncheck_clean %in% c("", "none", "-99", "i don't know") |
    !str_detect(attentioncheck_clean, "[0-9]") & 
    !str_detect(attentioncheck_clean, "two|doscientos") ~ "Missing",
  
  # incorrect
  TRUE ~ "Incorrect"
)

table(attentioncheck_table)
prop <- prop.table(table(attentioncheck_table))
```
The number of failures on the second attention check decreases from `r round(attcheck2.fails(data_raw$ATTCHECK_RES), 0)`% to `r round(prop.table(prop)["Incorrect"] * 100, 0)`%.

These are candidates of carless respondents and therefore "flagged" in this section.

## Para Data (Elnari Potgieter)

This section looks at the response time indices of the USA data set, considering total response time, average time per item, relative completion time and speeding factor.

1. Total Response Time

Total response time refers to the time it takes a respondent to complete the survey. The data set contains a "Duration" variable, analysed here, capturing response time in seconds. 

```{r}
# Key Descriptives of Duration variable (in seconds)
describe (data_raw$Duration, na.rm = TRUE)
```
Mean (1090.11) larger than median (620), showing right / positive skew in data. Minimum at 10 seconds, and maximum at 213592 seconds. Further analysis to look out for very high value outliers.

Further consideration to percentiles / quartiles.

```{r}
# Consider quartiles
skim(data_raw$Duration)
# Shows missing value, so proceed and exclude missing values
# Exclude NA values and skim the Duration variable
skim(data_raw$Duration[!is.na(data_raw$Duration)])
```
Maximum/p100 (213592) much higher than mean/p50 (620) and p75 (1101). The extreme high outliers may be due to respondents opening the survey and only answering the survey later. This does not necessarily mean they were not paying attention when they were answering, so I don't think their answers are necessarily invalid. However, on the lower end, 10 seconds for a survey with 111 items (Mede et al., 2025) - some of which open ended - sounds very short. So too 283 for p25. The extreme outliers also impact further calculations.

Box plots (in seconds) for further investigation.

```{r}
# Boxplot excluding NA values
boxplot(data_raw$Duration[!is.na(data_raw$Duration)],
        main = "Boxplot of Durations",
        ylab = "Duration (seconds)",
        col = "lightblue",
        border = "darkblue")
```
Boxplot reflects very high value outliers, which is less of concern in terms of response quality than the very low value durations as the former may still have paid attention, but they do impact the mean and median, which impacts further analysis. I considered running further analysis without the 1st and99th percentile respondents, but after consulting the course example and Appendix B of [-@leiner2019], it appears that extreme outliers are not necessarily removed at this point. 

2. Average Time per Item

The average time per item indicator captures the average (mean) time spent on survey questions.

```{r}
# Number of items
num_items <- 111

# Calculate the average time per item, excluding NA values in Duration
average_time_per_item <- data_raw$Duration/ num_items

# Compute the mean of all average times per item, excluding NA values
mean_average_time_per_item <- mean(average_time_per_item, na.rm = TRUE)

# Output the results
cat("Average time per item for each respondent (in seconds):\n")
cat("\nMean of all average times per item (in seconds):", mean_average_time_per_item, "seconds\n")
```
Using the filtered durations, the average time per item is 9.82 seconds.

I do think it prudent to consider the low value outliers more. I proceed by looking at the number and proportion of respondents with an average time per item below 2 seconds (as per the course example).

```{r}
# Identify respondents who fall below the cutoff of 2 seconds (as per example in course)
cutoff <- 2
below_cutoff <- which(average_time_per_item < cutoff)

# Count of respondents below the cutoff
count_below_cutoff <- length(below_cutoff)

# Proportion of respondents below the cutoff
proportion_below_cutoff <- count_below_cutoff / length(data_raw$Duration)

cat("\nRespondents below the cutoff of", cutoff, "seconds:\n")
cat("\nCount of respondents below the cutoff:", count_below_cutoff, "\n")
cat("Proportion of respondents below the cutoff:", proportion_below_cutoff, "\n")
```
845 respondents (18,37%) had an average time per item below 2 seconds. I am reluctant to exclude respondents only based on this criteria and will proceed with further analysis.

3. Relative Time Completion

Relative time completion compares the speed of completing the questionnaire for a respondent to the median completion time of the sample. 

```{r}
# Median completion time of the sample
median_time <- 620
# As per descriptives

# Calculate the relative completion time for each respondent, excluding NA values
relative_completion_time <- data_raw$Duration / median_time

# Exclude NA values from the relative completion time
relative_completion_time <- relative_completion_time[!is.na(relative_completion_time)]

# Output the relative completion times
cat("Relative completion time for each respondent (excluding NA):\n")

# Summary statistics for relative completion times
cat("\nSummary of relative completion times:\n")
print(summary(relative_completion_time))
```
On the lower end, there are respondents much faster than the median. This aligns with earlier flagged concerns regarding some respondents' fast completion rates. 

As per the course example, we calculate how many respondents are faster than twice the median completion time.

```{r}
# Identify respondents who are faster than two times the median completion time
cutoff <- 2
faster_than_cutoff <- which(relative_completion_time > cutoff)

# Count of respondents faster than two times the median completion time
count_faster <- length(faster_than_cutoff)

# Proportion of respondents faster than two times the median completion time
proportion_faster <- count_faster / length(data_raw$Duration)

# Output the results
cat("\nCount of respondents faster than two times the median:", count_faster, "\n")
cat("Proportion of respondents faster than two times the median:", proportion_faster, "\n")
```
883 (20,24%) respondents where twice as fast as the median of the relative completion time. I am reluctant to exclude this many respondents based on this criteria, so I proceed with also considering the speeding factor as per the course example.


4. Speeding Factor

[@leiner2019] suggests a factor of larger than two as outliers.This is calculated using their response time in relation to the median of the sample. 

```{r}
# Median completion time of the sample
median_time <- 620

# Calculate the speeding factor for each respondent
speeding_factor <- median_time / data_raw$Duration

# Exclude NA values from the speeding factor
speeding_factor <- speeding_factor[!is.na(speeding_factor)]

# Output the speeding factors
cat("Speeding factor for each respondent (excluding NA):\n")

# Summary statistics for speeding factors (excluding NA values)
cat("\nSummary of speeding factors (excluding NA):\n")
print(summary(speeding_factor))
```
From the above summary of speeding factors, the data set does include respondents with a speeding factor above 2.

```{r}
# Identify outliers (speeding factor > 2)
outliers <- which(speeding_factor > 2)

# Count of outliers
outlier_count <- length(outliers)

# Proportion of outliers
outlier_proportion <- outlier_count / length(data_raw$Duration)

# Output the count and proportion of outliers
cat("\nCount of outliers (speeding factor > 2):", outlier_count, "\n")
cat("Proportion of outliers:", outlier_proportion, "\n")
```
The dataset has 1197 (27,44%) - more than a quarter of - respondents with speeding factor above 2.

5. Summary

With 845 respondents (18,37%) with an average time per item below 2 seconds, 883 (20,24%) respondents twice as fast as the median of the relative completion time and 1197 (27,44%) respondents with speeding factor above 2, it does appear like many respondents were answering the questionnaire quite speedily. 

I ran the same analysis, "trimming" the 1st and 99th percentile to see if the higher value outliers affect the results in a meaningful way, but this was not the case as the mean remains very similar. 

I would be reluctant to exclude respondents based on response times only, as 18-27% of responses would be lost this way. My suggestion would be to align respondents flagged for response times with other tests - for example attention checks, response patterns and response styles.


## Missing Data (YOUR NAME)

The third kind of data quality indicators are missing data. How much data is missing, which data missing pattern occurs and are there cases with many missing data? (in respect to your choosen instrument(s) / variable(s)).

```{r}



```



## Response Distributions (YOUR NAME)

The fourth kind of data quality indicators are response distributions. Which indicators do you deem relevant and which criterion do you use for flagging?

```{r}



```



## Response Styles (Josefina Sommer)

The last kind of data quality indicators are response styles. Which style(s) do you want to focus on and how do you flag concerning cases?

To enable coherent analyses of response styles (1.) and patterns (2.), the longest possible sequence of 5pt Likert scale items entailing our focal scale - the SCIPOP scale - was used. I deliberately chose not to harmonize these items, allowing the authentic capture of natural differences in scale use across items without introducing artefactual distortions through rescaling. This preserves the ordinal nature of the data as well as item-specific variance, which is essential for the precise identification of patterns.

### 1. Response Styles
```{r}

#load packages, if not installed, install first
packages <- c("resquin", "gt", "ggcorrplot")

for (pkg in packages) {
  if (!require(pkg, quietly = TRUE, character.only = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
    library(pkg, character.only = TRUE)
  }
}

# 1: Response styles
all5pt <- data_raw %>%
select(ID_QUALTRICS, starts_with("GOALS_PRIO_"),
starts_with("GOALS_TACKLE_"),
starts_with("NORMPERC_"),
starts_with("WILLVUL_"),
TRUST_METHOD,
TRUST_PEW,
starts_with("OUTSPOKEN_"),
starts_with("SCIPOP_"))

# Numeric IDs
all5pt <- all5pt %>%
mutate(ID_num = as.integer(row_number()))

all5pt_only <- all5pt %>%
select(starts_with("GOALS_PRIO_"),
starts_with("GOALS_TACKLE_"),
starts_with("NORMPERC_"),
starts_with("WILLVUL_"),
TRUST_METHOD,
TRUST_PEW,
starts_with("OUTSPOKEN_"),
starts_with("SCIPOP_")) %>%
mutate(across(everything(), ~ as.integer(as.numeric(.)))) # Safe Integer-Konversion

all5pt_RS <- all5pt_only %>%
  resp_styles(
    scale_min = 1,
    scale_max = 5,
    id = all5pt$ID_num
  )

# Missings?
all5pt_RS_missing <- all5pt_RS %>%
  select(-id) %>%       
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "Indicator",
               values_to = "Missing_Count") %>%
  mutate(
    Total = nrow(all5pt),                
    Missing_Percent = Missing_Count / Total * 100
  )

all5pt_RS_missing 

# rerun analyses with allowing 50% of items to be unanswered to check if there are less missings then
min_prop <- 0.5

all5pt_RS2 <- all5pt_only %>%
  resp_styles(
    scale_min = 1,
    scale_max = 5,
    id = all5pt$ID_num,
    min_valid_responses = min_prop
  )

all5pt_RS2_missing <- all5pt_RS2 %>%
  select(-id) %>%       
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "Indicator",
               values_to = "Missing_Count") %>%
  mutate(
    Total = nrow(all5pt),                
    Missing_Percent = Missing_Count / Total * 100
  )

all5pt_RS2_missing 

# does not change much (from 23.3% to 22.0% => decide to drop all of them)

```

23.3% of U.S. participants do not have response style indicators due to missing answers, as the minimum proportion of valid responses was set to 100%. Reducing this threshold (e.g., to 50%) only marginally decreased the proportion of missing response style indicators (to approximately 22%). Given this negligible gain in sample size, alongside the increased arbitrariness in threshold selection, I retained the more stringent criterion of complete responses and excluded all cases with any missing values.
Accordingly, only respondents who answered all 5-point items (GOALS_PRIO_, GOALS_TACKLE_, NORMPERC_, WILLVUL_, etc.) were included in the response style analyses, resulting in the exclusion of roughly 23% of the sample. The subsequent exploratory analyses of response styles and response patterns are therefore based on the remaining 3.562 respondents.

```{r}
all5pt_RS_clean <- all5pt_RS %>%
  drop_na()

# Summary
str(all5pt_RS_clean)
head(all5pt_RS_clean)

# Plot Response styles
all5pt_RS_clean %>%
  select(-id) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(value*100))+
  geom_histogram(fill = "steelblue", color = "white") +
  facet_wrap(vars(name))+
  labs(x = "% of respondents responses",
       y = "Count")+
  theme_classic(base_size = 15)
```

For ARS (tendency to agree), responses are spread across the full range, with moderate peaks at both low and high percentages, indicating that some respondents rarely agree uniformly, while others tend to agree frequently.

DRS (tendency to disagree) shows a strong concentration near 0%, meaning that very few respondents systematically disagree with items, while some show moderate use of this style.

ERS (extreme response style) also spans a wide range but has visible peaks toward 0% and 100%, suggesting that some respondents either avoid or predominantly use extreme endpoints.

MRS (midpoint response style) shows higher frequencies near 0% and 100%, indicating that many respondents either rarely or frequently choose the midpoint option.

NERS (non-extreme response style) similarly displays peaks toward both extremes, implying that respondents either rarely or frequently avoid extreme categories.

### 2. Response patterns
```{r}
# 2: Response patterns
all5pt_RP <- all5pt_only %>%
  resp_patterns(
    id = all5pt$ID_num
    )

str(all5pt_RP)
head(all5pt_RP)

# Missings?
 all5pt_RP %>%
     select(-id) %>%
   summarise(across(everything(), ~ sum(is.na(.)))) %>%
   pivot_longer(everything(),
               names_to = "Indicator",
                values_to = "Missing_Count") %>%
   mutate(
     Total = nrow(all5pt_RP),
     Missing_Percent = Missing_Count / Total * 100
   )
# Missing_patterns, drop
 all5pt_RP_clean <- all5pt_RP %>%
   drop_na() # see above
 
```

To flag potentially suspect respondents, I focused on two response pattern indicators: longest string length and number of transitions between different response options.
Longest string length captures over‑consistent response patterns, while the number of transitions indicate how much respondents switch between options.

``` {r}
# 3: Flagging

# Data‑driven flagging of potentially suspect respondents
all5pt_RP_clean <- all5pt_RP_clean %>%
  mutate(
    thresh_long = quantile(longest_string_length, 0.95),
    thresh_low_trans = quantile(n_transitions, 0.05),
    flag_suspect = as.numeric(
      longest_string_length >= thresh_long |
      n_transitions <= thresh_low_trans
    )
  )

thresholds <- all5pt_RP_clean %>%
  slice(1) %>%  
  select(thresh_long, thresh_low_trans)

thresholds

# Attach the flag to all5pt_RS_clean
all5pt_RS_clean <- all5pt_RS_clean %>%
  left_join(
    all5pt_RP_clean %>% select(id, flag_suspect),
    by = "id"
  )

# How many respondents are flagged as suspicious?
all5pt_RP_clean %>%
  summarise(
    n_total = n(),
    n_suspect = sum(flag_suspect),
    prop_suspect = mean(flag_suspect)
  )

# Plot Response patterns
all5pt_RP_clean %>%
  select(-id, -thresh_long, -thresh_low_trans) %>%
  pivot_longer(
    cols = c(n_transitions, longest_string_length),
    names_to = "indicator",
    values_to = "value"
  ) %>%
  mutate(
    indicator = case_when(
      indicator == "longest_string_length" ~ "Longest string",
      indicator == "n_transitions" ~ "Transitions",
      TRUE ~ indicator
    )
  ) %>%
  ggplot(aes(value, fill = factor(flag_suspect))) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(vars(indicator), scales = "free_x", nrow = 1) +
  scale_fill_manual(
    values = c("0" = "steelblue", "1" = "firebrick"),  
    labels = c("0" = "Not suspicious", "1" = "Suspicious"), 
    name = "Status" 
  ) +
  labs(x = "Indicator value", y = "Number of respondents") +
  theme_classic(base_size = 15)

```

For each indicator, the top 5% or bottom 5% of the distribution were defined as ‘extreme’.
A respondent was flagged as potentially suspect if at least one of these two indicators was in the respective extreme tail.
This approach emphasizes unusual response patterns rather than specific response styles (e.g., extreme or mid‑point responding) and may therefore be less confounded with substantive attitudes. I flagged potentially careless respondents complemented by low transition counts, capturing monotonous responding.

Respondents in the extreme 5% tails were marked as suspicious: top 5% longest string length OR bottom 5% transitions. 
This data-driven approach prioritizes response patterns over style indicators (ERS/MRS), reducing confounding with substantive attitudes.

Out of 3.562 complete cases, this identified 259 respondents (roughly 7%) as potentially careless.


## All together (ALL)

Now that you have different flags you may want to combine them. Which combinations do you deem suitable? And how to (partly) filter the identified cases?

```{r}



```

## Discuss (ALL)

Discuss central findings shortly! Make some statements on your final decision in respect to include/exclude rules (cutoffs/thresholds/filter combination), data loss, consequences for planned analyses, and any notable what you like to adress.




### Response Styles and Patterns
Response styles and pattern analyses based on a broad set of 5-point Likert items revealed substantial heterogeneity in scale use across respondents. To ensure reliable estimation, analyses were restricted to complete cases, resulting in the exclusion of approximately 23% of the sample; lowering the minimum response threshold did not meaningfully reduce data loss.

While response style indicators (e.g., agreement, midpoint, and extreme responding) showed considerable variation, potentially suspect cases were primarily identified using response pattern indicators. Specifically, respondents were flagged based on extreme values in longest response strings and low numbers of response transitions, capturing overly consistent or monotonous answering behavior.

Using this conservative, data-driven approach, approximately 7% of the 3.562 respondents were identified as potentially careless. By focusing on response patterns rather than style indicators, this strategy reduces confounding with substantive attitudes and complements other data quality measures such as attention checks and paradata.


## References (ALL)

::: {#refs}
:::

::: {.callout-important appearance="default"}

## Pay Attention Please!
For final submission upload your prepared Quarto document and an exported html-file 
into Moodle, week 8, section "Final Assignment Course D2 Survey LU8.1".
**FINAL DEADLINE will be June 20th, 2025, sharp!**

:::

<span style="color:#1E8CC8;">**Good luck and see you next Thursday (June 5th) on our last weekly touchpoint!**</span>👋



